---
title: "Clean and Prepare Test/Calibration Data"
output: 
  html_document:
    theme: cosmo
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
date: 2018-06-27
author: Rich Pauloo
---

The purpose to this script is to clean and prepare the test data for model calibration in `06_calibration.Rmd`.

```{r, echo = FALSE, warning = FALSE, message=FALSE}
library(knitr)
opts_chunk$set(
  fig.width  = 7,
  fig.height = 7,
  collapse   = TRUE,
  message = FALSE,
  error = FALSE,
  warning = FALSE,
  cache = TRUE
)
```

Packages used.
```{r}
library(lubridate) # for dates
library(here)      # for system agnostic file paths
library(raster)    # for raster objects
library(tidyverse) # general purpose data science toolkit
library(sp)        # spatial objects
library(tibbletime)# for tibble time grouped summaries
```


Load points, CV shapefile, and ensure everything is in the correct projection.
```{r}
# mercator projection
merc <- crs("+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")

# read points and transform to mercator
dw <- shapefile(here("data","dry_wells", "DrywellReports_20180523_v1_100thDegPrecision.shp"))
dw <- spTransform(dw, merc)

# read central valley alluvial basin boundary and transform to mercator
cv <- shapefile(here("data","spatial","central_valley_alluvial_boundary",
                     "Alluvial_Bnd.shp"))
cv <- spTransform(cv, merc)
```

First glimpse of the data.
```{r}
plot(cv)
points(dw, pch = 19, cex = 0.1)
```

Subset for points within the CV.
```{r}
dwcv <- dw[cv, ]
```

Ensure they plot in the CV.
```{r}
plot(cv)
points(dwcv, pch = 19, cex = .1)
```

Subset the points to domestic well outages during the 2012-2016 drought. First need to merge spatial data with another table containing this information.
```{r}
# read in data without geometry (points)
dw_dat <- readxl::read_xlsx(here("data","dry_wells","DrywellReports_20180523_v1_PersonalInfoRedacted.xlsx"))

# replace all cases of "groundwater shortage" with the same ID
dw_dat$Shortages <- 
  dw_dat %>% 
    pull(Shortages) %>% 
    str_replace(c("Pump not working","pump stopped working.",
                  "lack of water, well level dropping- intermitent water production"), 
                  "Dry well (groundwater)")

# clean dates
dw_dat <- dw_dat %>% 
  rename(issue_date = `Approximate Issue Start Date`,    # better column name
         report_date = `Record Creation Date`)           # better column name

dw_dat$issue_date  <- lubridate::ymd(dw_dat$issue_date)  # POSIXct to Date
dw_dat$report_date <- as_date(dw_dat$report_date)        # POSIXct to Date

dw_dat <- dw_dat %>% 
  # if the issue date is missing, replace with report date
  mutate(issue_date = ifelse(is.na(issue_date), report_date, issue_date)) %>% 
  mutate(issue_date = as_date(issue_date)) # put back into date format

# subset to events during the 2012-2016 drought
dw_dat2 <- dw_dat %>% 
  filter(year(issue_date) >= 2012 & year(issue_date) <= 2016) %>% 
  rename(Drywell_ID = `Drywell ID`)

# perform the join of tabluar and spatial data
dwcv@data$Drywell_ID <- as.numeric(dwcv@data$Drywell_ID) # convert to numeric for join
dwcv@data <- left_join(dwcv@data, dw_dat2, by = "Drywell_ID") %>% # join by ID
  mutate(issue_date = ifelse(is.na(issue_date),   # if issue dates are missing
                             as_date(Report_Dat), # give them report dates
                             issue_date), 
         issue_date = as_date(issue_date))  

# filter out dates from spatial object that don't match drought years
dwcv1 <- dwcv[which(year(dwcv@data$issue_date) >= 2012 & 
                    year(dwcv@data$issue_date) <= 2016), ]  
```

Take a look at dry well reports over time. 
```{r, include = FALSE}
dwtt <- as_tbl_time(dwcv1@data, index = issue_date)
dwtt %>% 
  arrange(issue_date) %>% 
  collapse_by("monthly") %>% 
  group_by(issue_date) %>% 
  summarise(count = n ()) %>% 
  ggplot() +
  geom_line(aes(issue_date, count), lwd = 1) +
  geom_rect(aes(xmin = make_date(2012, 01, 01), xmax = make_date(2016, 12, 31),
                ymin = 0, ymax = 1150), fill = "red", alpha = 0.01) +
  coord_cartesian(ylim = c(0, 1100)) +
  scale_y_continuous(breaks = seq(0,1000,250), limits = seq(0,100,250)) +
  labs(title = "Domestic Well Failure Reports",
       subtitle = "Aggregated By Month",
       y = "Count", x = "Date") +
  theme_minimal()
```

Export for validation script.
```{r}
shapefile(dwcv1, here("data", "dry_wells", "cleaned_dw", "dwcv.shp"))
```

