---
title: "Domestic Well Failure Model Calibration 2"
output: 
  html_document:
    theme: cosmo
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
    highlight: "pygments"
date: 2018-07-03
author: Rich Pauloo
---

The purpose of this script is to generate a distribution of domestic well retirement ages from the OSWCR database and with the help of bootstrapping.

```{r, echo = FALSE, warning = FALSE, message=FALSE}
library(knitr)
opts_chunk$set(
  fig.width  = 7,
  fig.height = 7,
  collapse   = TRUE,
  message    = FALSE,
  error      = FALSE,
  warning    = FALSE,
  cache      = TRUE
)
```


`dat.rds` was generated from `publish_3.Rmd`. The only difference between it and `clean_dat`, is that it represents all domestic wells throughout the state, and there is no culling because of lack of latitude/longitude data (around 2% of total data).  

```{r}
library(tidyverse)
library(sp)

dat <- read_rds("dat.rds")
```


# Entire State of California

Find duplicate domestic wells, and calculate a distribution of retirement ages.
```{r}
# all legacy log numbers
l <- dat %>% filter(type == "unused"& !is.na(LegacyLogNumber)) %>% pull(LegacyLogNumber)

# all legacy log numbers with two entries
two_entries <- dat %>% 
  filter(LegacyLogNumber %in% l & 
         !is.na(LegacyLogNumber) &
         !is.na(DateWorkEnded)) %>% 
  count(LegacyLogNumber) %>% 
  arrange(desc(n)) %>% 
  filter(n >= 2) %>% 
  pull(LegacyLogNumber)

# another filter
tef <- dat %>% 
  filter(LegacyLogNumber %in% two_entries) %>% 
  dplyr::select(LegacyLogNumber, DateWorkEnded, type) %>% 
  filter(type %in% c("domestic", "unused")) %>% 
  count(LegacyLogNumber) %>%
  filter(n >= 2) %>% 
  pull(LegacyLogNumber)

# 
d2 <- dat %>% 
  filter(LegacyLogNumber %in% tef & !is.na(DateWorkEnded)) %>% 
  dplyr::select(LegacyLogNumber, DateWorkEnded, type) %>% 
  arrange(LegacyLogNumber) %>% 
  tidyr::separate(DateWorkEnded, into = c("month","day","year")) %>% 
  mutate(year = as.numeric(year)) %>% 
  dplyr::select(LegacyLogNumber, year, type) 
  
# split into list
dl <- base::split(d2, f = d2$LegacyLogNumber)

# function to get retirement age
get_ra <- function(x){
  y1 = min(x$year)
  y2 = max(x$year)
  return(abs(y1-y2))
}

# retirement age distribution
rad <- sapply(dl, get_ra)

```


```{r}
# remove impossible values
rad <- rad[rad >=1 ]

# bootstrap with replacement
sample(rad, size = 50000, replace = T) %>% hist()

# fit lognormal dist to data
library(fitdistrplus)

lnd <- fitdistrplus::fitdist(rad, # to water_above_bot values
                             distr = "lnorm",    # lognormal distribution
                             method = "mle")     # use maximum liklihood

plot(lnd)

m <- lnd$estimate[1] # mean
v <- (lnd$estimate[2])^2 # variance

# calculate the mean of the lognormal dist
mu <- exp(m + .5*(v))

# extract fitted `density` values given the paramaters from the fit 
dlnd <- dlnorm(seq(1,70,.1), coef(lnd)[1], coef(lnd)[2])
  
# bind the paramaaters in a dataframe
dlnd <- data.frame(x = seq(1,70,.1), d = dlnd)
  
# plot
dom_ret_age <- data.frame(rad=rad) %>% 
  ggplot() +
  geom_histogram(aes(x = rad, y = ..density..), binwidth = 3, color = "white") + 
  geom_line(data = dlnd, aes(x, d), color = "red", lwd = 1) +
  geom_vline(xintercept = mu, linetype = "dashed", lwd = 1) + 
  geom_text(aes(19, 0.06, label = "mean = 16")) + 
  theme_minimal() +
  labs(x = "Age (years)", y = "Probability Density",
       title = "Domestic Well Retirement Age in California", 
       subtitle = "(nwells = 188)")

dom_ret_age
write_rds(dom_ret_age, "dom_ret_age.rds")
```

***  

# Central Valley

Same, but for `dat2df`
```{r}
# find data in cv
cv <- read_rds("cv.rds")

dat2 <- dat %>% filter(!is.na(lon) & !is.na(lat))

pts <- cbind(dat2$lon, dat2$lat)

dat2sp <- SpatialPointsDataFrame(pts, 
                                 data = dat2, 
                                 proj4string = 
                                   crs("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")) 

cvll <- spTransform(cv, crs("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

# crop dat2sp to cv
dat2c <- dat2sp[cvll, ]

# get just data
dat2df <- dat2c@data


#################################

# all legacy log numbers
l <- dat2df %>% filter(type == "unused"& !is.na(LegacyLogNumber)) %>% pull(LegacyLogNumber)

# all legacy log numbers with two entries
two_entries <- dat2df %>% 
  filter(LegacyLogNumber %in% l & 
         !is.na(LegacyLogNumber) &
         !is.na(DateWorkEnded)) %>% 
  count(LegacyLogNumber) %>% 
  arrange(desc(n)) %>% 
  filter(n >= 2) %>% 
  pull(LegacyLogNumber)

# another filter
tef <- dat2df %>% 
  filter(LegacyLogNumber %in% two_entries) %>% 
  select(LegacyLogNumber, DateWorkEnded, type) %>% 
  filter(type %in% c("domestic", "unused")) %>% 
  count(LegacyLogNumber) %>%
  filter(n >= 2) %>% 
  pull(LegacyLogNumber)

# 
d2 <- dat2df %>% 
  filter(LegacyLogNumber %in% tef & !is.na(DateWorkEnded)) %>% 
  select(LegacyLogNumber, DateWorkEnded, type) %>% 
  arrange(LegacyLogNumber) %>% 
  tidyr::separate(DateWorkEnded, into = c("month","day","year")) %>% 
  mutate(year = as.numeric(year)) %>% 
  select(LegacyLogNumber, year, type) 
  
# split into list
dl <- base::split(d2, f = d2$LegacyLogNumber)

# function to get retirement age
get_ra <- function(x){
  y1 = min(x$year)
  y2 = max(x$year)
  return(abs(y1-y2))
}

# retirement age distribution
rad <- sapply(dl, get_ra)

```


```{r}
# remove impossible values
rad <- rad[rad >=1 ]

# bootstrap with replacement
sample(rad, size = 10000, replace = T) %>% hist()


# fit lognormal dist to data
library(fitdistrplus)

lnd <- fitdistrplus::fitdist(rad, # to water_above_bot values
                             distr = "lnorm",    # lognormal distribution
                             method = "mle")     # use maximum liklihood

plot(lnd)

m <- lnd$estimate[1] # mean
v <- (lnd$estimate[2])^2 # variance

# calculate the mean
exp(m + .5*(v))
```





