---
title: "Domestic Well Failure Model Calibration"
output: 
  html_document:
    theme: cosmo
    toc: TRUE
    toc_float: TRUE
    toc_depth: 2
    code_folding: hide
    highlight: "pygments"
date: 2018-07-23
author: Rich Pauloo
---

# Introduction

After a conversation with Herve, this script attemps to quantify and propogate uncertainty thoughout the model. In particular I focus on four areas of uncertainty:  

* **Uncertainty in domestic well retirement age**: addressed by calculating a distribution of domestic well retirement ages from the OSWCR (`05_distribution_retirement_age.Rmd`), and using the mean retirement age (16 years) to determine what wells from OSWCR are included in the model.  
* **Uncertainty in well failure reporting**: addressed by calculating the distribution of failure proportions given well failure data from the Department of Water Resources, and selecting validation townships based on this distribution. Townships in the 2nd and 3rd quartiles of the distribution serve as calibration targets.  
* **Uncertainty in OSWCR reporting**: addressed by generating the distribution of well completion reports per township, and taking the all but the 1st quantile. Townships are likely to have less than 100% completion rates, and it is impossible to have a completion rate greater than 100%.  
* **Uncertainty in maximum groundwater level during the 2012-2016 drought**: addressed by introducing a tuning parameter (water level scaling factor) that adjusts the interpolated groundwater level values within an optimization problem.  

This script uses the cleaned test data of well failure during the 2012-2016 drought (`dwcv.shp`) to compute 3 models of increasing complexity: (1) a null model, (2) a single parameter optimization, and lastly (3) a two-parameter optimization. The tuning parameters in the model are $d$ & $\omega$.  

$d$ is the proportion of the distance between the land surface and the bottom of the screened interval that the pump sits at. For example, if the pump sits halfway between the land surface and the bottom of the screened interval, $d = 0.5$. Thus, $0 \le d \le 1$, and a $d$ near $1$, indicates the pump is near the bottom of the screened interval.  

$\omega$ is water level scaling factor. The parameter is constrained by the 5th and 95th quantiles of the distribution of minimum and maximum groundwater elevations during the 2012-2016 drought. For example, during the drought, the minimum and maximum value for each raster cell form a minimum and maximum layer. The values from this minimum and maximum layer form distributions of groundwater levels, and the 5th and 95th quantiles form the lower and upper bounds of the allowable water levels, thus constraining $\omega$ within a range of feasible values.  


***  

```{r, echo = FALSE, warning = FALSE, message=FALSE}
library(knitr)
opts_chunk$set(
  fig.width  = 7,
  fig.height = 7,
  collapse   = TRUE,
  message    = FALSE,
  error      = FALSE,
  warning    = FALSE,
  cache      = TRUE
)
```

# Data Setup

Packages used.
```{r}
library(here)      # for system agnostic file paths
library(GISTools)  # counting points in polygons with poly.counts()
library(raster)    # for raster objects
library(tidyverse) # general purpose data science toolkit
library(sp)        # spatial objects
```

Load Central Valley, domestic well failures (test set), and active domestic well (train set) shapefiles.
```{r}
# mercator projection
merc <- crs("+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 
            +ellps=WGS84 +datum=WGS84 +units=m +no_defs")

# read central valley alluvial basin boundary and transform to mercator
cv <- shapefile(here("data","spatial","central_valley_alluvial_boundary",
                     "Alluvial_Bnd.shp"))
cv <- spTransform(cv, merc)

# townships shapefile in central valley
blmcv <- read_rds("blmcv.rds")

# gsa shapefile in central valley
gsacv <- read_rds("gsacv.rds")

# bulletin 118 shapefile in central valley
b118cv <- read_rds("b118cv.rds")

# read cleaned test data and transform to mercator
dw <- shapefile(here("data", "dry_wells", "cleaned_dw", "dwcv.shp"))
dw <- spTransform(dw, merc)

# read the training data
domcv6 <- read_rds("domcv6_mean_gw.rds")

# read in water level data
ml <- read_rds("ml.rds")

# baseline water level defined by the spring measurements
#baseline <- ml[[1]]$Ensemble

# with spplot
# spplot(baseline, 
#        col.regions = rev(get_col_regions()),
#        main = "Spring 2011 Groundwater Depth Below Land Surface")

# with ggplot
# as.data.frame(baseline, xy = TRUE) %>% 
#   ggplot(aes(x,y)) +
#   geom_raster(aes(x,y, fill = Ensemble)) +
#   coord_fixed(1.1) + 
#   theme_void() +
#   labs(fill = "Feet",
#        title = "Groundwater Depth Below Land Surface",
#        subtitle = "Spring 2011") +
#   scale_fill_continuous(type = "viridis", na.value="transparent") 
```

First we calculate the column height between land surface and each of these points, which is simply the depth to the bottom of the screened interval.
```{r}
# We use the Spring 2011 groundwater level as the baseline water column height, 
# and extract these data to the spatial points.
# domcv6 <- raster::extract(baseline, domcv6, # from baseline, get values @pts
#                           # extracted value = avgerage of 4 nearest cells
#                           method = "bilinear", 
#                           # add values onto data.frame of sp object
#                           sp = TRUE)           

# calculate column height from bottom of screened interval 
# and land surface
domcv6@data$wch_2 <- domcv6@data$bot - 0

# remove wells that are already dry, about 2% of data
domcv6 <- domcv6[which(domcv6@data$wch_2 >= 0), ]

# re-transform
domcv6 <- spTransform(domcv6, merc)

# make a copy for the optimization
domcv8 <- domcv7 <- domcv6
```

Visualize where we have data on well failures.
```{r}
# see where we have testing data
plot(b118cv)
points(dw, pch = 19, cex = 0.5, col = "blue")
```

***  

# Uncertainty 

Here we quantify and propogate uncertainty in domestic well retirement age, well failure reporting, and OSWCR reporting. In the calibration itself, we account for uncertainty in interpolated groundwater levels during the drought.  

## Domestic Well Retirement Age 

We calculate a distribution of domestic well retirement ages from the OSWCR in a separate script (`05_distribution_retirement_age.Rmd`). The mean retirement age (16 years) is not used to determine what wells from OSWCR are included in the model because it is simply too low to be believable. 

```{r}
dom_ret_age <- read_rds("dom_ret_age.rds")
dom_ret_age
```

Anecdotally, it is not uncommon for wells to retire at 30-50 years, and a model run with 16 years as the retirement age systematically underpredicts well failure rates in calibration townships, and the total number of wells that fail during the 2012-2016 drought on the whole.  

In another script (`06_calibration_herve_alvar.Rmd`), the domestic well retirement age is calibrated by iterating over retirement ages from 10-60 years. The relationship between retirement age, calibration SSE, pump submergence, and the total number of dry wells predicted by the model using that retirement age are shown. A retirement age of **33 years**, with the second lowest SSE is selected. The count of dry wells predicted with this retirement age is almost exactly the number of reported dry wells. The reason why the count of dry wells does not monotonically increase with retirement age is because this output is conditional on the pump submergence, which is calibrated, and affects the number of wells that go dry for a particular retirement age.  

The next steps in the calibration were written into an inner loop in `06_calibration_herve_alvar.Rmd` with the calibration age iterating on the outside. These steps are shown to illustrate how uncertainy was quantified and propogated through the model.  

```{r}
ret_age_df <- read_rds("ret_age_df.rds") # df of retirement ages and associated metrics
d_opt <- ret_age_df$d[24]

p_grid <- read_rds("calib_p_grid.rds")
p_grid
```


## Well Failure Reporting

We calculate the distribution of failure proportions given well failure data from the Department of Water Resources and wells in the OSWCR with a retirement age of 33 years. Validation townships are selected based on this distribution. Townships in the 2nd and 3rd quartiles of the distribution serve as calibration targets. Low failure rates can be attributed to underreporting, and failure rates above 1, and high failure rates in general imply underreporting from the OSWCR database. The red shaded area shows the townships kept in the validation set.  

```{r}
# retirement age
ret_age <- 33 # mean of dist = 16.  `06_calibration_herve_alvar` = 33 yrs

cutoff <- 2012 - ret_age

awells <- domcv6 # active wells
awells <- awells[awells@data$year >= cutoff, ] # subset out retrired wells


# count up the observations in each township
dw_counts <- poly.counts(dw, blmcv)
wr_counts <- poly.counts(awells, blmcv)

# failure ratio
fr <- dw_counts / wr_counts
fr2 <- fr[fr > 0 & !is.na(fr) & !is.nan(fr)]

# find quartiles
quart <- quantile(fr2, c(0.25, 0.75))

# histogram
data.frame(fr = fr2) %>% ggplot(aes(fr)) + 
  geom_histogram(binwidth = 0.05, col = "white") + 
  geom_vline(xintercept = c(quart[1], quart[2]), col = "red") +
  geom_rect(aes(xmin = quart[1], xmax = quart[2], ymin = 0, ymax = 48), 
            fill = "red", alpha = 0.002) +
  theme_minimal() +
  labs(title = "Domestic Well Failure Rates",
       subtitle = "Central Valley Townships: 2012-2016 Drought",
       x = "Failure Rate", y = "Count") +
  # zoom in on the main trend, disregarding 3 wells at 3,4,5
  coord_cartesian(xlim = c(0,1)) 


# can't have failure ratio greater than 1
fr[is.na(fr) & is.nan(fr)] <- 0
calibration_poly <- blmcv[fr >= quart[1] & fr <= quart[2], ]

# extract failure rates for calibration polygons
calibration_fr <- fr[fr >= quart[1] & fr <= quart[2]]

# dry adn wet wells for calibration set to observe
calibration_dry <- dw_counts[fr >= quart[1] & fr <= quart[2] & !is.nan(fr)]
calibration_wet <- wr_counts[fr >= quart[1] & fr <= quart[2] & !is.nan(fr)]
```


```{r, echo = FALSE}
# Visualize
#plot(calibration_poly, main = "Calibration Townships & Observations")
#points(dw[calibration_poly, ], pch = 19, cex = 0.5, col = "blue")

calibration_poly$test <- calibration_fr
#spplot(calibration_poly, "test", main = "Failure Rates at Calibration Townships")
```

## OSWCR Reporting

Another source of uncertainty is that wells in the OSWCR are likely to exhibit underreporting. This will inflate the observed and predicted failure rate. By culling validation townships with high failure rates in the step above, the problem is partially addressed. To add another layer of complexity, we calculate the distribution of well counts in our calibration townships, and select townships from the 2nd, 3rd, and 4th quartiles to use in the calibration to not let under-reported townships influence the calibration. The red shaded area shows the townships kept in the validation set.  
```{r}
# count oswcr wells in polygons
oswcr_in_calibration <- poly.counts(awells, calibration_poly)

# find quartiles
quart2 <- quantile(oswcr_in_calibration, c(0.25, 1))

# histogram
data.frame(x = oswcr_in_calibration) %>% 
  ggplot(aes(oswcr_in_calibration)) + 
  geom_histogram(binwidth = 10, col = "white") + 
  geom_vline(xintercept = c(quart2[1], quart2[2]), col = "red") +
  geom_rect(aes(xmin = quart2[1], xmax = quart2[2], ymin = 0, ymax = 15), 
            fill = "red", alpha = 0.002) +
  theme_minimal() +
  labs(title = "Count of Domestic Wells per Township",
       subtitle = "Calibration Townships: 1996-2012",
       x = "Number of Domestic Wells", y = "Count") 

# subset the calibration polygons to remove low OSWCR count polygons
calibration_poly2 <- calibration_poly[oswcr_in_calibration >= quart2[1] & oswcr_in_calibration <= quart2[2], ]
# subset the calibration failure rate vector to remove low OSWCR count polygons
calibration_fr2 <- calibration_fr[oswcr_in_calibration >= quart2[1] & oswcr_in_calibration <= quart2[2]]

# remove 1 troublesome township with unreasoably high predicted failure
calibration_poly2 <- calibration_poly2[-40, ]
calibration_fr2 <- calibration_fr2[-40]
```

Final polygons used in the validation set. 
```{r}
plot(calibration_poly2, main = "Validation Townships/Observations")
points(dw[calibration_poly2, ], pch = 19, cex = 0.5, col = "blue")

spplot(calibration_poly2, "test", main = "Failure Rate")
```

## Water Level

Uncertainty in water level interpolations between 2012-2016 are addressed in **Modeling: Double Parameter Optimization**.   


***  

# Modeling

Three models are built, each one increasing in complexity, and obtaining a better fit to the validation data. The three models to follow are:  

1. **null model**: the most simple possible model used as a benchmark to evaluate other models  
2. **single parameter optimization** (decision variable: $d$)  
3. **two-parameter optimization** (decision variables: $d$ & $\omega$)  

## Null Model


First we calculate a null model of well failure. No calibration is involved, and we call a well "dry" simply if the mean water level during the 20120-2016 drought falls below the screen bottom.
```{r}
# well is dry if max negative gw level falls at 
# or below bottom of perforated interval
domcv7 <- domcv7[domcv7@data$year >= cutoff, ] # subset for active wells
domcv7@data <- domcv7@data %>% 
  mutate(dry = ifelse(max_gw >= bot, TRUE, FALSE))

# count dry and wet wells in calibration townships

# dry wells
town_dry <- poly.counts(domcv7[domcv7@data$dry == T, ], calibration_poly2) 

# wet wells
town_wet <- poly.counts(domcv7[domcv7@data$dry == F, ], calibration_poly2) 

# predicted failure ratios: should be less than actual because in the model
# wells go dry when water falls at or below the screened interval bottom
# and wells should fail before this because pumps sit well above this level
frp <- town_dry / (town_dry + town_wet)

# calculate error
# loss function (SSE): de-emphasizes penalty for tails
SSE <- function(observed, predicted) {
  (sum((predicted - observed)^2, na.rm=TRUE))
}

SSE(frp, calibration_fr2)

# visualize
data.frame(obs = calibration_fr2, pred = frp) %>% 
  ggplot() +
  geom_point(aes(obs, pred)) +
  geom_text(aes(obs, pred, label = names(calibration_fr2)), hjust = 1.2) + 
  geom_abline(slope = 1) +
  coord_cartesian(ylim = c(0,.2), xlim = c(0,.2)) + 
  theme_minimal() +
  labs(title = "Null Model",
       subtitle = "Observed v. Predicted Failure Rates (validation townships = 39)",
       x = "Observed", y = "Predicted")
```

The simple model actually does pretty well, but seems to underpredict most townships.  


## Single Parameter Optimization

We saw that the simple model wasn't good enough to predict well failure, so let's now try a more complicated model that adjusts for the pump's position in the well. If the water level falls below that, then the well fails.  

We previously calibrated the pump separation proportion, `d` with single parameter optimization and `optimize()`. The optimal value was found to be $d =$ `r round(d_opt, 2)`.
```{r}
# calculate
d <- d_opt

domcv9 <- domcv8[which(domcv8@data$year >= cutoff), ] 

# water column heights
wch <- domcv9@data$bot -0
  
# pump separation distance (ft) from top of water column
ps <- d * wch
  
# pump location (ft below land surface)
pl <- 0 + ps
  
# put into dataframe and compute well failures
domcv9@data <- domcv9@data %>% 
  mutate(pl = pl,
          dry = ifelse(max_gw >= pl, TRUE, FALSE))

# count dry and wet wells in calibration townships
town_dry <- poly.counts(domcv9[domcv9@data$dry == T, ], calibration_poly2)#dry
town_wet <- poly.counts(domcv9[domcv9@data$dry == F, ], calibration_poly2)#wet
  
# predicted failure ratios
frp2 <- town_dry / (town_dry + town_wet)

# SSE
SSE(frp2, calibration_fr2)

# visualize
data.frame(obs = calibration_fr2, pred = frp2) %>% 
  ggplot() +
  geom_point(aes(obs, pred)) +
  geom_text(aes(obs, pred, label = names(calibration_fr2)), hjust = 1.2) + 
  geom_abline(slope = 1) +
  coord_cartesian(ylim = c(0,.2), xlim = c(0,.2)) +
  theme_minimal() +
  labs(title = "Single Parameter Optimization",
       subtitle = "Observed v. Predicted Failure Rates (validation townships = 39)",
       x = "Observed", y = "Predicted")
```

Calibration of this single parameter improves the SSE by about 50%. 


## Double Parameter Optimization

Can we improve the model by calibrating the water levels that go into the simulation?  

Now we try calibrating $d$ and $\omega$, where $\omega$ is the water level scaling factor.  

First we need to find possible values of $\omega$. We constrain the water scaling factor to be a value that transforms the mean groundwater level, without resulting in a distribution of water levels that exceed the 90th percentile of the maximum water levels, nor falls below the 10th percentile of the minimum water levels. That is:  

```{r}
# raster stack of 2012-2016 drought GW levels
es <- read_rds("es.rds") 
es <- projectRaster(es, crs = merc)

# constrain the water level by the 10% quantiles of the min and max
q10 <- log(min(es)@data@values) %>% quantile(., 0.05, na.rm = T)
q90 <- log(max(es)@data@values) %>% quantile(., 0.95, na.rm = T)

# solve for the upper and lowr bounds via grid search
wmin = vector(length = 200)
wmax = vector(length = 200)
step = seq(0.01,2, 0.01)
for(i in 1:200){
  wmin[i] <- log((mean(es) * step[i])@data@values) %>% quantile(., 0.05, na.rm = T)
  wmax[i] <- log((mean(es) * step[i])@data@values) %>% quantile(., 0.95, na.rm = T)
}

# find the upper and lower limits on the water scaling factor
li <- wmin[wmin >= q10] %>% min() # lower boundary index
ui <- wmax[wmax <= q90] %>% max() # upper boundary index

wl <- step[which(wmin == li)] # lower limit on w (water level scaling factor)
wu <- step[which(wmax == ui)] # upper limit on w 
```

Now with solved constraints, $0.47 \le \omega \le 1.36$, we can proceed with two parameter optimization.  
```{r}
# read in data
domcv <- read_rds("domcv.rds") # unaltered domestic wells in cv - shapefile
domcv@data$year <- as.numeric(domcv@data$year) # get dates into numeric class
domcv <- spTransform(domcv, merc)

# function to optimize
fun <- function(x, w) {
  d <- x[1]                 # optimize: proportion of WCH distance
  w <- x[2]                 # optimize: water level scale factor
  if (d >= 1)   return(Inf) # constraint: d cannot be greater than 1
  if (d < .001) return(Inf) # constraint: d cannot be less than 0
  if (w >= wu)  return(Inf) # constraint: w cannot be greater than wu
  if (w <= wl)  return(Inf) # constraint: w cannot be less than wl
  
  # apply water level scaling factor and extract level for each of point
  emax <- mean(es) * w # mean groundwater depth below land surface 
  names(emax) <- "max_gw" # change the name of the raster layer

  # extract the raster max gw level values and bind to spatial points dataframe
  domcv <- raster::extract(emax,     # extract from emax
                           domcv,    # to domcv points
                           sp = TRUE)# and add the values to a data frame

  # remove the ~1000 wells that fall outside of the raster cells
  domcv <- domcv[!is.na(domcv@data$max_gw), ] 
  
  # copy of the data without missing values and imposisble dates
  domcv2 <- domcv[which(!is.na(domcv@data$year) & 
                        domcv@data$year <= 2017 &  # remove impossible values
                        domcv@data$year >= 1900), ]# out of range
  
  # ry <- 2012 - a            
  ry <- cutoff
  
  # subset for active wells
  domcv3 <- domcv2[which(domcv2@data$year >= ry), ] 
  
  # water column heights
  # negative water column heights do not affect optimization
  # these are just wells that are very far gone
  wch <- domcv3@data$bot - 0
  
  # pump separation distance (ft) from top of water column
  ps <- d * wch
  
  # pump location (ft below land surface)
  pl <- 0 + ps
  
  # put into dataframe and compute well failures
  domcv3@data <- domcv3@data %>% 
    mutate(pl = pl,
           dry = ifelse(max_gw >= pl, TRUE, FALSE))

  # count dry and wet wells in calibration townships
  town_dry <- poly.counts(domcv3[domcv3@data$dry == T, ], calibration_poly2)#dry
  town_wet <- poly.counts(domcv3[domcv3@data$dry == F, ], calibration_poly2)#wet
  
  # predicted failure ratios
  frp <- town_dry / (town_dry + town_wet)
  
  # calculate error
  return(SSE(calibration_fr2, frp))

}

# optimize with sep distance and retirement age
opt <- optim(c(.8, 1.2),  # initial paramater values
             fn = fun) 
opt
```

## Calibration Results 

Calibration of water levels within reasonable bounds given the data we have gives us more reduction in the loss function, but not considerably so. We now view the calibration results with our calibrated parameters for $d$ and $\omega$.    
```{r}
# run the optimized parameters to get predicitons

  # apply water level scaling factor and extract level for each of point
  emax <- mean(es) * opt$par[2] # mean groundwater depth below land surface 
  names(emax) <- "max_gw" # change the name of the raster layer

  # extract the raster max gw level values and bind to spatial points dataframe
  domcv2 <- raster::extract(emax,     # extract from emax
                            domcv,    # to domcv points
                            sp = TRUE)# and add the values to a data frame

  # remove the ~1000 wells that fall outside of the raster cells
  domcv2 <- domcv2[!is.na(domcv2@data$max_gw), ] 
  
  # copy of the data without missing values and imposisble dates
  domcv3 <- domcv2[which(!is.na(domcv2@data$year) & 
                        domcv2@data$year <= 2017 &  # remove impossible values
                        domcv2@data$year >= 1900), ]# out of range
  
  # ry <- 2012 - a            
  ry <- cutoff
  
  # subset for active wells
  domcv4 <- domcv3[which(domcv3@data$year >= ry), ] 
  
  # water column heights
  # negative water column heights do not affect optimization
  # these are just wells that are very far gone
  wch <- domcv4@data$bot - 0
  
  # pump separation distance (ft) from top of water column
  ps <- opt$par[1] * wch
  
  # pump location (ft below land surface)
  pl <- 0 + ps
  
  # put into dataframe and compute well failures
  domcv4@data <- domcv4@data %>% 
    mutate(pl = pl,
           dry = ifelse(max_gw >= pl, TRUE, FALSE))

  # count dry and wet wells in calibration townships
  town_dry <- poly.counts(domcv4[domcv4@data$dry == T, ], calibration_poly2)#dry
  town_wet <- poly.counts(domcv4[domcv4@data$dry == F, ], calibration_poly2)#wet

  
# predicted failure ratios
frp3 <- town_dry / (town_dry + town_wet)

# SSE
SSE(frp3, calibration_fr2)

# visualize
data.frame(obs = calibration_fr2, pred = frp3) %>% 
  ggplot() +
  geom_point(aes(obs, pred)) +
  geom_text(aes(obs, pred, label = names(calibration_fr2), hjust = 1.2)) +
  geom_abline(slope = 1) +
  coord_cartesian(ylim = c(0,.2), xlim = c(0,.2)) +
  theme_minimal() +
  labs(title = "Double Parameter Optimization ",
       subtitle = "Observed v. Predicted Failure Rates (validation townships = 39)",
       x = "Observed", y = "Predicted")
```

## Comparing Models 

We've developed 3 models of well failure. In the null model (1), we assumed a maximum depth of groundwater equal to the mean of the groundwater levels during the 2012-2016 drought. In the single parameter optimization (2), we tuned a parameter $d$, which positioned the pump around 75% between land surface and the well screen bottom. In the final two-parameter optimization model (3), in addition to re-tuning $d$, which was found to be around 85%, we tuned $\omega$, or the water level scaling factor, allowing for water levels greater than the mean (which we expect), but less than the maximum observed groundwater level.  

With increasing model complexity, initial gains in model accuracy are obtained, but moving past single parameter optimization does not yield much reduction in SSE. This highlights the importance of the position of the well pump as an important parameter in the model.  

```{r}
data.frame(null = SSE(frp, calibration_fr2),
           single_param = SSE(frp2, calibration_fr2),
           double_param = SSE(frp3, calibration_fr2)) %>% 
  gather(model, SSE) %>% 
  mutate(SSE = round(SSE,3)) %>% 
  kable()
```


## Extending the Calibrated Model to Make Predictions

Now we take the calibrated parameters to the rest of the dataset, and compute well failure at the Township, GSA, and Bulletin 118 subbasin level.
```{r}
# run the optimized parameters to get predicitons

  # apply water level scaling factor and extract level for each of point
  emax2 <- mean(es) * opt$par[2] # mean groundwater depth below land surface 
  names(emax2) <- "max_gw" # change the name of the raster layer

  # extract the raster max gw level values and bind to spatial points dataframe
  domcv5 <- raster::extract(emax2,     # extract from emax
                            domcv,    # to domcv points
                            sp = TRUE)# and add the values to a data frame

  # remove the ~1000 wells that fall outside of the raster cells
  domcv5 <- domcv5[!is.na(domcv5@data$max_gw), ] 
  
  # copy of the data without missing values and imposisble dates
  domcv5 <- domcv5[which(!is.na(domcv5@data$year) & 
                        domcv5@data$year <= 2017 &  # remove impossible values
                        domcv5@data$year >= 1900), ]# out of range
  
  # ry <- 2012 - a            
  ry <- cutoff
  
  # subset for active wells
  domcv5 <- domcv5[which(domcv5@data$year >= ry), ] 
  
  # water column heights
  # negative water column heights do not affect optimization
  # these are just wells that are very far gone
  wch <- domcv5@data$bot - 0
  
  # pump separation distance (ft) from top of water column
  ps <- opt$par[1] * wch
  
  # pump location (ft below land surface)
  pl <- 0 + ps
  
  # put into dataframe and compute well failures
  domcv5@data <- domcv5@data %>% 
    mutate(pl = pl,
           dry = ifelse(max_gw >= pl, TRUE, FALSE))

# now predict well failure at the relevant scales

# GSAs
gsa_dry <- poly.counts(domcv5[domcv5$dry == T, ], gsacv)
gsa_wet <- poly.counts(domcv5[domcv5$dry == F, ], gsacv)
gsa_frp <- gsa_dry / (gsa_dry + gsa_wet)

gsacv@data$frp <- gsa_frp * 100
gsacv@data$fc <- paste0("(",gsa_dry,"/",(gsa_dry+gsa_wet)," = ",round(gsa_frp*100,2),"%)")
gsacv@data$dry <- gsa_dry
gsacv@data$wet <- gsa_wet

# B118 
b118_dry <- poly.counts(domcv5[domcv5$dry == T, ], b118cv)
b118_wet <- poly.counts(domcv5[domcv5$dry == F, ], b118cv)
b118_frp <- b118_dry / (b118_dry + b118_wet)

b118cv@data$frp <- b118_frp * 100
b118cv@data$fc <- paste0("(",b118_dry,"/",(b118_dry+b118_wet)," = ",round(b118_frp*100,2),"%)")
b118cv@data$dry <- b118_dry
b118cv@data$wet <- b118_wet

# Township
blm_dry <- poly.counts(domcv5[domcv5$dry == T, ], blmcv)
blm_wet <- poly.counts(domcv5[domcv5$dry == F, ], blmcv)
blm_frp <- blm_dry / (blm_dry + blm_wet)

blmcv@data$frp <- blm_frp * 100
blmcv@data$fc <- paste0("(",blm_dry,"/",(blm_dry+blm_wet)," = ",round(blm_frp*100,2),"%)")
blmcv@data$dry <- blm_dry
blmcv@data$wet <- blm_wet
```

***  

# Results

Visualize dry well counts and failure ratios at Township, GSA, and Bulletin 118 Subbasin scales with both static and interactive maps.  

## Static Maps

Predictions of domestic well failure for the 2012-2016 drought over different areal units.

```{r}
library(sf)
library(ggplot2)

# jet color palette
jet_colors <- colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan", "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))

# convert to sf
blmcvsf  <- st_as_sf(blmcv)
gsacvsf  <- st_as_sf(gsacv)
b118cvsf <- st_as_sf(b118cv)
cvsf     <- st_as_sf(cv)

# plot townships
p1 <- ggplot() +
  geom_sf(data = blmcvsf, aes(fill = frp)) +
  geom_sf(data = cvsf, alpha = 0.01, lwd = .8, color = "red") +
  scale_fill_viridis_c("% Failure") +
  coord_sf(crs = st_crs(102003)) +
  labs(title = "Failure Ratio",
       y = "Latitude", x = "Longitude") +
  theme_bw()

p1c <- ggplot() +
  geom_sf(data = blmcvsf, aes(fill = dry)) +
  geom_sf(data = cvsf, alpha = 0.01, lwd = .8, color = "red") +
  scale_fill_gradientn(colors = jet_colors(7)) +
  coord_sf(crs = st_crs(102003)) +
  labs(title = "Failure Count",
       y = "Latitude", x = "Longitude") +
  theme_bw() + 
  labs(fill = "Count")

# plot GSAs
p2 <- ggplot() +
  geom_sf(data = gsacvsf, aes(fill = frp)) +
  geom_sf(data = cvsf, alpha = 0.01, lwd = .8, color = "red") +
  scale_fill_viridis_c("% Failure") +
  coord_sf(crs = st_crs(102003)) +
  labs(title = "Failure Ratio",
       y = "Latitude", x = "Longitude") +
  theme_bw()

p2c <- ggplot() +
  geom_sf(data = gsacvsf, aes(fill = dry)) +
  geom_sf(data = cvsf, alpha = 0.01, lwd = .8, color = "red") +
  scale_fill_gradientn(colors = jet_colors(7)) +
  coord_sf(crs = st_crs(102003)) +
  labs(title =    "Failure Count",
       y = "Latitude", x = "Longitude") +
  theme_bw() + 
  labs(fill = "Count")

# plot Bulltein 118 subbasins
p3 <- ggplot() +
  geom_sf(data = b118cvsf, aes(fill = frp)) +
  geom_sf(data = cvsf, alpha = 0.01, lwd = .8, color = "red") +
  scale_fill_viridis_c("% Failure") +
  coord_sf(crs = st_crs(102003)) +
  labs(title = "Failure Ratio",
       y = "Latitude", x = "Longitude") +
  theme_bw()

p3c <- ggplot() +
  geom_sf(data = b118cvsf, aes(fill = dry)) +
  geom_sf(data = cvsf, alpha = 0.01, lwd = .8, color = "red") +
  scale_fill_gradientn(colors = jet_colors(7)) +
  coord_sf(crs = st_crs(102003)) +
  labs(title =    "Failure Count",
       y = "Latitude", x = "Longitude") +
  theme_bw() + 
  labs(fill = "Count")

# save
# ggsave(p1, file = "p_blm.png", dpi = 300, height = 10, width = 7)
```

### Township
```{r}
library(cowplot)
plot_grid(p1c, p1, align = "h")
```

***  

### Groundwater Sustainability Agency
```{r}
plot_grid(p2c, p2, align = "h")
```

***  

### Bulletin 118 Subbasin
```{r}
plot_grid(p3c, p3, align = "h")
```


***  

## Interactive Maps

Predictions for the 2012-2016 drought.

```{r}
library(leaflet)
library(colormap)

# transform polygons to lat/lon for leaflet
b118cvsf <- b118cvsf %>% st_transform(crs = "+init=epsg:4326") 
gsacvsf  <- gsacvsf  %>% st_transform(crs = "+init=epsg:4326") 
blmcvsf  <- blmcvsf  %>% st_transform(crs = "+init=epsg:4326") 

# transform points to ll for leaflet
domcv5ll <- spTransform(domcv5, crs("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0") )

# round well bottom measurements
domcv5ll$bot <- round(domcv5ll$bot, 2)

# icons for points
icons <- icons(
  
)

# function to get colors
getColor <- function(d) {
  sapply(d$dry, function(dry) {
    ifelse(dry == TRUE, "red", "blue") # red if fail, blue if not
  })
}

# make marker list
icons <- awesomeIcons(
  icon = 'f041',
  iconColor = 'black',
  library = 'fa',
  markerColor = getColor(domcv5ll@data)
)


```


### Township
```{r}
pal <- colorBin(palette = colormap(colormaps$viridis, nshades = 10),
                domain = blmcvsf$frp, bins = seq(0,100,10))

pal2 <- colorBin(palette = colormap(colormaps$jet, nshades = 90),
                 domain = blmcvsf$dry, bins = seq(0,90,10))

# center coordinates for `setView` function
clat <- st_bbox(blmcvsf)[c(2,4)] %>% mean()
clng <- st_bbox(blmcvsf)[c(1,3)] %>% mean()

blmcvsf %>% 
  # only show townships with 30 or more wells to begin with
  #mutate(frp = ifelse(nwells >= 30, frp, NA)) %>% 
  leaflet(width = "100%") %>% 
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(label = ~ paste(TWNSHPLAB, fc),
              # polygons
              fillColor = ~ pal2(dry), 
              fillOpacity = 0.7, 
              smoothFactor = 0.5,
              group = "Count of Dry Wells",
              # lines
              stroke = TRUE, 
              color = "#323232", 
              opacity = 1, 
              weight = 1) %>% 
  addPolygons(label = ~ paste(TWNSHPLAB, fc),
              # polygons
              fillColor = ~ pal(frp), 
              fillOpacity = 0.7, 
              smoothFactor = 0.5,
              group = "Failure Ratio",
              # lines
              stroke = TRUE, 
              color = "#323232", 
              opacity = 1, 
              weight = 1) %>% 
  addAwesomeMarkers(lng = domcv5ll@coords[, 1],
             lat = domcv5ll@coords[, 2],
             popup = paste("Well ID:", domcv5ll$WCRNumber,"<br>",
                           "(", domcv5ll$lon, "N", domcv5ll$lat, "W)", "<br>",
                           "Bottom of Perforated Interval:", domcv5ll$bot, "ft.", "<br>",
                           "Dry:", domcv5ll$dry),
             icon = icons,
             group = "Wells",
             clusterOptions = markerClusterOptions()) %>%
  addLegend("bottomright", 
            pal = pal2, 
            values = ~ dry,
            opacity = 1,
            title = "Count of Dry Wells",
            group = "Count of Dry Wells",
            labFormat = function(type, cuts, p) {
              n = length(cuts)
              paste0(cuts[-n], " &ndash; ", cuts[-1])
            }) %>% 
  addLegend("topright", 
            pal = pal, 
            values = ~ frp,
            opacity = 1,
            title = "% Failure",
            group = "Failure Ratio",
            labFormat = function(type, cuts, p) {
              n = length(cuts)
              paste0(cuts[-n], " &ndash; ", cuts[-1])
            }) %>% 
  addLayersControl(overlayGroups = c("Count of Dry Wells", "Failure Ratio", "Wells"), 
                   position = "topleft",
                   options = layersControlOptions(collapsed = FALSE)) %>% 
  hideGroup(c("Failure Ratio","Wells")) %>% 
  setView(lat = clat, lng = clng, zoom=7) %>% 
  addEasyButton(easyButton(
    icon="fa-globe", title="Zoom to Level 7",
    onClick=JS("function(btn, map){ map.setZoom(7); }"))) %>%
  addEasyButton(easyButton(
    icon="fa-crosshairs", title="Locate Me",
    onClick=JS("function(btn, map){ map.locate({setView: true}); }")))

```

***  

### Groundwater Sustainability Agency
```{r}
pal <- colorBin(palette = colormap(colormaps$viridis, nshades = 10),
                domain = gsacvsf$frp, bins = seq(0,100,10))

pal2 <- colorBin(palette = colormap(colormaps$jet, nshades = 10),
                 domain = gsacvsf$dry, bins = seq(0,200,20))

gsacvsf %>% 
  leaflet(width = "100%") %>% 
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(label = ~ paste(as.character(GSA.Name), fc),
              # polygons
              fillColor = ~ pal2(dry), 
              fillOpacity = 0.7, 
              smoothFactor = 1,
              group = "Count of Dry Wells",
              # lines
              stroke = TRUE, 
              color = "#323232", 
              opacity = 1, 
              weight = 1) %>% 
  addPolygons(label = ~ paste(as.character(GSA.Name), fc),
              # polygons
              fillColor = ~ pal(frp), 
              fillOpacity = 0.7, 
              smoothFactor = 1,
              group = "Failure Ratio",
              # lines
              stroke = TRUE, 
              color = "#323232", 
              opacity = 1, 
              weight = 1) %>% 
  addAwesomeMarkers(lng = domcv5ll@coords[, 1],
             lat = domcv5ll@coords[, 2],
             popup = paste("Well ID:", domcv5ll$WCRNumber,"<br>",
                           "(", domcv5ll$lon, "N", domcv5ll$lat, "W)", "<br>",
                           "Bottom of Perforated Interval:", domcv5ll$bot, "ft.", "<br>",
                           "Dry:", domcv5ll$dry),
             icon = icons,
             group = "Wells",
             clusterOptions = markerClusterOptions()) %>%
  hideGroup("Wells") %>%
  addLegend("topright", 
            pal = pal, 
            values = ~ frp,
            opacity = 1,
            title = "% Failure",
            group = "Failure Ratio",
            labFormat = function(type, cuts, p) {
              n = length(cuts)
              paste0(cuts[-n], " &ndash; ", cuts[-1])
            }
            ) %>%
  addLegend("bottomright", 
            pal = pal2, 
            values = ~ dry,
            opacity = 1,
            title = "Count of Dry Wells",
            group = "Count of Dry Wells",
            labFormat = function(type, cuts, p) {
              n = length(cuts)
              paste0(cuts[-n], " &ndash; ", cuts[-1])
            }
            ) %>%
  addLayersControl(overlayGroups = c("Count of Dry Wells", "Failure Ratio", "Wells"), 
                   position = "topleft",
                   options = layersControlOptions(collapsed = FALSE)) %>% 
  hideGroup(c("Failure Ratio","Wells")) %>% 
  addEasyButton(easyButton(
    icon="fa-globe", title="Zoom to Level 7",
    onClick=JS("function(btn, map){ map.setZoom(7); }"))) %>%
  addEasyButton(easyButton(
    icon="fa-crosshairs", title="Locate Me",
    onClick=JS("function(btn, map){ map.locate({setView: true}); }")))
```

***  

### Bulletin 118 Subbasin
```{r}
pal <- colorBin(palette = colormap(colormaps$viridis, nshades = 10),
                domain = b118cvsf$frp, bins = seq(0,100,10))

pal2 <- colorBin(palette = colormap(colormaps$jet, nshades = 10),
                 domain = b118cvsf$dry, bins = seq(0,300,30))

b118cvsf %>% 
  leaflet(width = "100%") %>% 
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(label = ~ paste(as.character(Subbasin_N), fc),
              # polygons
              fillColor = ~ pal2(dry), 
              fillOpacity = 0.7, 
              smoothFactor = 1,
              group = "Count of Dry Wells",
              # lines
              stroke = TRUE, 
              color = "#323232", 
              opacity = 1, 
              weight = 1) %>% 
  addPolygons(label = ~ paste(as.character(Subbasin_N), fc),
              # polygons
              fillColor = ~ pal(frp), 
              fillOpacity = 0.7, 
              smoothFactor = 1,
              group = "Failure Ratio",
              # lines
              stroke = TRUE, 
              color = "#323232", 
              opacity = 1, 
              weight = 1) %>% 
  addAwesomeMarkers(lng = domcv5ll@coords[, 1],
             lat = domcv5ll@coords[, 2],
             popup = paste("Well ID:", domcv5ll$WCRNumber,"<br>",
                           "(", domcv5ll$lon, "N", domcv5ll$lat, "W)", "<br>",
                           "Bottom of Perforated Interval:", domcv5ll$bot, "ft.", "<br>",
                           "Dry:", domcv5ll$dry),
             icon = icons,
             group = "Wells",
             clusterOptions = markerClusterOptions()) %>%
  addLegend("topright", 
            pal = pal, 
            values = ~ frp,
            opacity = 1,
            title = "% Failure",
            group = "Failure Ratio",
            labFormat = function(type, cuts, p) {
              n = length(cuts)
              paste0(cuts[-n], " &ndash; ", cuts[-1])
            }
            ) %>% 
  addLegend("bottomright", 
            pal = pal2, 
            values = ~ dry,
            opacity = 1,
            title = "Count of Dry Wells",
            group = "Count of Dry Wells",
            labFormat = function(type, cuts, p) {
              n = length(cuts)
              paste0(cuts[-n], " &ndash; ", cuts[-1])
            }
            ) %>% 
  addLayersControl(overlayGroups = c("Count of Dry Wells", "Failure Ratio", "Wells"), 
                   position = "topleft", 
                   options = layersControlOptions(collapsed = FALSE)) %>% 
  hideGroup(c("Failure Ratio","Wells")) %>% 
  addEasyButton(easyButton(
    icon="fa-globe", title="Zoom to Level 7",
    onClick=JS("function(btn, map){ map.setZoom(7); }"))) %>%
  addEasyButton(easyButton(
    icon="fa-crosshairs", title="Locate Me",
    onClick=JS("function(btn, map){ map.locate({setView: true}); }")))

```

