---
title: "Domestic Well Failure Model Calibration"
output: 
  html_document:
    theme: cosmo
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
    highlight: "pygments"
date: 2018-07-03
author: Rich Pauloo
---

# Introduction

The purpose to this script use the cleaned test data of well failure during the 2012-2016 drought (`dwcv.shp`) to calibrate two tuning parameters ($d$ & $a$) in the physical model.  

$d$ is the proportion of total water column height that the pump sits in. For example, if the pump sits halfway between the top of the water column and the bottom of the screened interval, $d = 0.5$. Formally, $0 \le d \le 1$, and for $d$ near $1$, the pump is near the bottom of the screened interval.  

$a$ is the age at which wells are retired. Average reasonable values for this parameter range from $20-60$ years, so these are set as constraints on retirement age: $20 \le d \le 60$.  

***  

# Assumptions

This study assumes that in townships with well failure reports $\ge 60$, that the reporting is compelete for the township. This allows us to calculate a failure ratio for townships with failures $\ge 60$ during the drought, and we further assume that calibration townships represet a spatially unbiased sample of townships, so that the calibrated parameters can be used across the Central Valley. We also assume that the count of OSCWR domestic well reports in a calibration township represents the total number of domestic wells in that township. The OSWCR database may underrepresent the number of wells because some wells are unreported, leading to an underestimation of the total number of domestic wells in a township. Morevoer, utilizing the full record of wells in the database also overestimates the number of domestic wells in a township because some proportion of the wells are bound to be retired. The underreporting rate and true retirement age are difficult to constrain, but they bias the results in different directions and so have the effect of acting as checks on the other, rather than systematically biasing results in one direction. 

***  

# Code

```{r, echo = FALSE, warning = FALSE, message=FALSE}
library(knitr)
opts_chunk$set(
  fig.width  = 7,
  fig.height = 7,
  collapse   = TRUE,
  message    = FALSE,
  error      = FALSE,
  warning    = FALSE,
  cache      = TRUE
)
```

## Data Setup

Packages used.
```{r}
library(here)      # for system agnostic file paths
library(raster)    # for raster objects
library(tidyverse) # general purpose data science toolkit
library(sp)        # spatial objects
library(GISTools)  # counting points in polygons with poly.counts()
```

Load Central Valley, domestic well failures (test set), and active domestic well (train set) shapefiles.
```{r}
# mercator projection
merc <- crs("+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 
            +ellps=WGS84 +datum=WGS84 +units=m +no_defs")

# read central valley alluvial basin boundary and transform to mercator
cv <- shapefile(here("data","spatial","central_valley_alluvial_boundary",
                     "Alluvial_Bnd.shp"))
cv <- spTransform(cv, merc)

# townships shapefile in central valley
blmcv <- read_rds("blmcv.rds")

# gsa shapefile in central valley
gsacv <- read_rds("gsacv.rds")

# bulletin 118 shapefile in central valley
b118cv <- read_rds("b118cv.rds")

# read cleaned test data and transform to mercator
dw <- shapefile(here("data", "dry_wells", "cleaned_dw", "dwcv.shp"))
dw <- spTransform(dw, merc)

# read the training data
domcv6 <- read_rds("domcv6.rds")

# read in water level data
ml <- read_rds("ml.rds")

# baseline water level defined by the spring measurements
baseline <- ml[[1]]$Ensemble

# with spplot
# spplot(baseline, 
#        col.regions = rev(get_col_regions()),
#        main = "Spring 2011 Groundwater Depth Below Land Surface")

# with ggplot
as.data.frame(baseline, xy = TRUE) %>% 
  ggplot(aes(x,y)) +
  geom_raster(aes(x,y, fill = Ensemble)) +
  coord_fixed(1.1) + 
  theme_void() +
  labs(fill = "Feet",
       title = "Groundwater Depth Below Land Surface",
       subtitle = "Spring 2011") +
  scale_fill_continuous(type = "viridis", na.value="transparent") 
```

Now we predict well failure ratio at the level that we have data from. Optimize a tuning parameter $d$.

We use the Spring 2011 groundwater level as the baseline water column height, and extract these data to the spatial points. 
```{r}
domcv6 <- raster::extract(baseline, domcv6, # from baseline, get values @pts
                          # extracted value = avgerage of 4 nearest cells
                          method = "bilinear", 
                          # add values onto data.frame of sp object
                          sp = TRUE)           

# calculate water column height from bottom of screened interval 
# and ensemble prediction
domcv6@data$wch_2 <- domcv6@data$bot - domcv6@data$Ensemble

# remove wells that are already dry, about 2% of data
domcv6 <- domcv6[which(domcv6@data$wch_2 >= 0), ]

# re-transform
domcv6 <- spTransform(domcv6, merc)

# make a copy for the optimization
domcv8 <- domcv7 <- domcv6
```

Calculate proportion of well failures at township level for townships with sufficient data. This will act as the calibration target. 
```{r}
# see where we have testing data
plot(b118cv)
points(dw, pch = 19, cex = 0.5, col = "blue")
```

***  

## Calibration

Let's take a look at the townships where we perform the calibration. Assume townships with 60 or more samples have 100% reporting, and that OSWCR wells in the calibration townships (no filtering on retirement age) represent true count of wells in that township.
```{r}
# count up the observations in each township
blmcv_counts <- poly.counts(dw, blmcv)

# those with 60 or more observations for the test set
# subset for the test set & verify that it aligns with test well locations
calibration_poly <- blmcv[blmcv_counts >= 60, ]
# plot the calibration townships
#plot(calibration_poly, main = "Calibration Townships & Observations") 
#points(dw, pch = 19, cex = 0.5, col = "blue")

# calculate failure ratio
fr <- poly.counts(dw, calibration_poly) / 
      poly.counts(domcv6, calibration_poly)

# can't have failure ratio greater than 1
calibration_poly <- calibration_poly[fr < 1, ]

fr <- fr[-7] # remove last observation

plot(calibration_poly, main = "Calibration Townships & Observations")
points(dw[calibration_poly, ], pch = 19, cex = 0.5, col = "blue")
```

Now we calculate well failure.
```{r}
# well is dry if max negative gw level falls at 
# or below bottom of perforated interval
domcv6@data <- domcv6@data %>% 
  mutate(dry = ifelse(max_gw >= bot, TRUE, FALSE))

# count dry and wet wells in calibration townships

# dry wells
town_dry <- poly.counts(domcv6[domcv6@data$dry == T, ], calibration_poly) 

# wet wells
town_wet <- poly.counts(domcv6[domcv6@data$dry == F, ], calibration_poly) 

# predicted failure ratios: should be less than actual because in the model
# wells go dry when water falls at or below the screened interval bottom
# and wells should fail before this because pumps sit well above this level
frp <- town_dry / (town_dry + town_wet)

# calculate error
# loss function (RMSE): de-emphasizes penalty for tails
RMSE <- function(observed, predicted) {
  sqrt(mean((predicted - observed)^2, na.rm=TRUE))
}
```

<!-- Now we can optimize. We seek to minimize the RMSE via a tuning parameter, $d$. -->
<!-- ```{r} -->
<!-- # function to optimize -->
<!-- fun <- function(x) { -->
<!--   d <- x                    # paramater to optimize: proportion of WCH distance -->
<!--   if (d >= 1)   return(Inf) # constraint: d cannot be greater than 1  -->
<!--   if (d < .001) return(Inf) # constraint: d cannot be less than 0 -->

<!--   # water column heights -->
<!--   # negative water column heights do not affect optimization -->
<!--   # these are just wells that are very far gone -->
<!--   wch <- domcv7@data$bot - domcv7@data$Ensemble -->

<!--   # pump separation distance (ft) from top of water column -->
<!--   ps <- d * wch -->

<!--   # pump location (ft below land surface) -->
<!--   pl <- domcv7@data$Ensemble + ps -->

<!--   # put into dataframe and compute well failures -->
<!--   domcv7@data <- domcv7@data %>%  -->
<!--     mutate(pl = pl, -->
<!--            dry = ifelse(max_gw >= pl, TRUE, FALSE)) -->

<!--   # count dry and wet wells in calibration townships -->
<!--   town_dry <- poly.counts(domcv7[domcv7@data$dry == T, ], calibration_poly) # dry -->
<!--   town_wet <- poly.counts(domcv7[domcv7@data$dry == F, ], calibration_poly) # wet -->

<!--   # predicted failure ratios -->
<!--   frp <- town_dry / (town_dry + town_wet) -->

<!--   # calculate error -->
<!--   return(RMSE(fr, frp)) -->
<!-- } -->


<!-- # single parameter optimization -->
<!-- set.seed(20150518) -->
<!-- opt <- optimize(fun,               # function to minimize -->
<!--                 interval = c(0,1), # interval to search -->
<!--                 tol = 0.001, -->
<!--                 maximum = FALSE)                     -->
<!-- opt  -->
<!-- ``` -->

<!-- Visualize the objective function space. -->
<!-- ```{r, eval=FALSE, include=FALSE} -->
<!-- # function to optimize -->
<!-- fun <- function(x) { -->
<!--   d <- x[1]                 # paramater to optimize: proportion of WCH distance -->
<!--   perc <- x[2] # percentage of repsondants -->
<!--   if (d >= 1)   return(Inf) # constraint: d cannot be greater than 1  -->
<!--   if (d < .001) return(Inf) # constraint: d cannot be less than 0 -->
<!--   if (perc >= 1)   return(Inf)  -->
<!--   if (perc < .001) return(Inf)  -->

<!--   respf <- 1/perc # respondent factor -->

<!--   # water column heights -->
<!--   # negative water column heights do not affect optimization -->
<!--   # these are just wells that are very far gone -->
<!--   wch <- domcv8@data$bot - domcv8@data$Ensemble -->

<!--   # pump separation distance (ft) from top of water column -->
<!--   ps <- d * wch -->

<!--   # pump location (ft below land surface) -->
<!--   pl <- domcv8@data$Ensemble + ps -->

<!--   # put into dataframe and compute well failures -->
<!--   domcv8@data <- domcv8@data %>%  -->
<!--     mutate(pl = pl, -->
<!--            dry = ifelse(max_gw >= pl, TRUE, FALSE)) -->

<!--   # count dry and wet wells in calibration townships -->
<!--   town_dry <- poly.counts(domcv8[domcv8@data$dry == T, ], calibration_poly) # dry -->
<!--   town_wet <- poly.counts(domcv8[domcv8@data$dry == F, ], calibration_poly) # wet -->

<!--   # predicted failure ratios -->
<!--   frp <- town_dry / (town_dry + town_wet) -->
<!--   frp <- frp * respf -->

<!--   # calculate error -->
<!--   return(RMSE(fr, frp)) -->
<!-- } -->

<!-- # optimize with sep distance and percentage reporting -->
<!-- opt <- optim(c(.5,.5),          # initial paramater values -->
<!--              fn = fun)  -->
<!-- opt -->

<!-- mat <- matrix(NA, nrow=99, ncol=99) -->
<!-- ii <- seq(.1,.3,.1) -->
<!-- for(i in 1:99){ -->
<!--   for(j in 1:99){ -->
<!--     temp <- optim(c(ii[i],ii[j]), fn = fun) -->
<!--     mat[i, j] = temp$value[1]  -->
<!--   } -->
<!-- } -->


<!-- # with vectorize -->
<!-- res <- outer( -->
<!--   1:3, -->
<!--   1:3, -->
<!--   Vectorize(function(i,j) { optim(c(ii[i],ii[j]), fn = fun)} ) -->
<!-- ) -->

<!-- raster(mat) %>% plot() -->
<!-- ``` -->

## Optimal Parameters

We calibrate the pump separation proportion, `d` and the welll `a`. 
```{r}
# function to optimize
# function to optimize
fun <- function(x) {
  d <- x[1]                 # optimize: proportion of WCH distance
  a <- x[2]                 # optimize: retirement age
  if (d >= 1)   return(Inf) # constraint: d cannot be greater than 1
  if (d < .001) return(Inf) # constraint: d cannot be less than 0
  if (a >= 60)   return(Inf)# upper bound on reitement age
  if (a < 20) return(Inf)   # lower bound on retirement age
  
  ry <- 2017 - a            # retirement year
  
  # subset for active wells
  domcv9 <- domcv8[which(domcv8@data$year >= ry), ] 
  
  # water column heights
  # negative water column heights do not affect optimization
  # these are just wells that are very far gone
  wch <- domcv9@data$bot - domcv9@data$Ensemble
  
  # pump separation distance (ft) from top of water column
  ps <- d * wch
  
  # pump location (ft below land surface)
  pl <- domcv9@data$Ensemble + ps
  
  # put into dataframe and compute well failures
  domcv9@data <- domcv9@data %>% 
    mutate(pl = pl,
           dry = ifelse(max_gw >= pl, TRUE, FALSE))

  # count dry and wet wells in calibration townships
  town_dry <- poly.counts(domcv9[domcv9@data$dry == T, ], calibration_poly)#dry
  town_wet <- poly.counts(domcv9[domcv9@data$dry == F, ], calibration_poly)#wet
  
  # actual failure ratios
  # calculate actual failure ratio
  # fr <- poly.counts(dw, calibration_poly) / 
  #       poly.counts(domcv9, calibration_poly)
  #fr[fr >= 1] <- 1

  # predicted failure ratios
  frp <- town_dry / (town_dry + town_wet)
  
  # calculate error
  return(RMSE(fr, frp))
}


# optimize with sep distance and retirement age
opt <- optim(c(.5,30),  # initial paramater values
             fn = fun) 
opt
```

## Null Model 

We also calculate a null model which assumes the calibrated retirement age, but $d = 1$, which is equivalent to the most simple model: that the pump sits at the well bottom. The null model returns more than 2 times the error of our optimzed model.
```{r}
# calculate null model
a = opt$par[2]
d = 1

########################################################################
  ry <- 2017 - a # retirement year

  # subset for active wells
  domcv9 <- domcv8[which(domcv8@data$year >= ry), ] 
  
  # water column heights
  # negative water column heights do not affect optimization
  # these are just wells that are very far gone
  wch <- domcv9@data$bot - domcv9@data$Ensemble
  
  # pump separation distance (ft) from top of water column
  ps <- d * wch
  
  # pump location (ft below land surface)
  pl <- domcv9@data$Ensemble + ps
  
  # put into dataframe and compute well failures
  domcv9@data <- domcv9@data %>% 
    mutate(pl = pl,
           dry = ifelse(max_gw >= pl, TRUE, FALSE))

  # count dry and wet wells in calibration townships
  town_dry <- poly.counts(domcv9[domcv9@data$dry == T, ], calibration_poly)#dry
  town_wet <- poly.counts(domcv9[domcv9@data$dry == F, ], calibration_poly)#wet
  
  # actual failure ratios
  # calculate actual failure ratio
  # fr <- poly.counts(dw, calibration_poly) / 
  #       poly.counts(domcv9, calibration_poly)
  # fr[fr >= 1] <- 1

  # predicted failure ratios
  frp <- town_dry / (town_dry + town_wet)
  
  # calculate error
  RMSE(fr, frp)

########################################################################
```

## Calibration Results 

We view the observated versus predicted failure ratios from our calibration. Overall, these calibration results are not bad.
```{r}
# optimal parameters
a = opt$par[2]
d = opt$par[1]

########################################################################
  ry <- 2017 - a # retirement year
  domcv9 <- domcv8[which(domcv8@data$year >= ry), ] # subset for active wells
  
  # water column heights
  # negative water column heights do not affect optimization
  # these are just wells that are very far gone
  wch <- domcv9@data$bot - domcv9@data$Ensemble
  
  # pump separation distance (ft) from top of water column
  ps <- d * wch
  
  # pump location (ft below land surface)
  pl <- domcv9@data$Ensemble + ps
  
  # put into dataframe and compute well failures
  domcv9@data <- domcv9@data %>% 
    mutate(pl = pl,
           dry = ifelse(max_gw >= pl, TRUE, FALSE))

  # count dry and wet wells in calibration townships
  town_dry <- poly.counts(domcv9[domcv9@data$dry == T, ], calibration_poly)#dry
  town_wet <- poly.counts(domcv9[domcv9@data$dry == F, ], calibration_poly)#wet
  
  # actual failure ratios
  # calculate actual failure ratio
  # fr <- poly.counts(dw, calibration_poly) / 
  #       poly.counts(domcv9, calibration_poly)
  #fr[fr >= 1] <- 1

  # predicted failure ratios
  frp <- town_dry / (town_dry + town_wet)
  
########################################################################  

# bind observed and predicted failure ratios to calibration townships
calibration_poly@data$fr  <- fr
calibration_poly@data$frp <- frp

#spplot(calibration_poly, c("fr","frp"))

# plot observed v predicted per township
data.frame(township = names(fr), fr = fr, frp = frp) %>% 
  ggplot(aes(fr, frp)) +
  geom_point() +
  geom_text(aes(label = township), nudge_y = -0.035, nudge_x = -0.01) +
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) +
  geom_abline(y=1, linetype = "dashed") +
  labs(title = "Calibrated Well Failure Proportions at Townships",
       subtitle = "Townships with reports >= 60",
       x = "Actual Failure", y = "Predicted Failure") +
  theme_minimal() -> p_obj

p_obj

#ggsave(p_obj, filename = "p_obj.png", dpi = 300, height = 8, width = 8)
```

<!-- And lastly, we visualize the objective function space to ensure we're not missing other minima in the optimization, or stuck in a local minima. -->
<!-- ```{r, eval = FALSE} -->
<!-- # visualize objective function space -->
<!-- obj_mat <- matrix(NA, nrow=25, ncol=25)  -->
<!-- ii <- seq(.01,.99,.04) -->
<!-- jj <- seq(25, 49, 1) -->
<!-- for(i in 1:25){ -->
<!--   for(j in 1:25){  -->
<!--     temp <- optim(c(ii[i],jj[j]), fn = fun)  -->
<!--     obj_mat[i, j] = temp$value[1]   -->
<!--   }  -->
<!-- }  -->

<!-- write_rds(obj_mat, "obj_mat.rds") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- obj_mat <- read_rds("obj_mat.rds") -->

<!-- obj_mat <- obj_mat[1:25,1:25] -->

<!-- raster(obj_mat) %>% plot() -->

<!-- raster(obj_mat) %>%  -->
<!--   as.data.frame(xy = TRUE) %>%  -->
<!--   mutate(x = rep(ii, 25), y = rep(rev(jj), each=25)) %>%  -->
<!--   ggplot(aes(x, y)) + -->
<!--   geom_raster(aes(x,y, fill = layer)) + -->
<!--   scale_fill_viridis_c() -->
<!-- ``` -->

## Apply the Calibrated Model

Apply the calibrated parameters to the rest of the dataset, and compute well failure at the Township, GSA, and Bulletin 118 subbasin level.
```{r}
opt_dist <- opt$par[1] # calibrated separation distance
opt_age  <- opt$par[2] # calibrated retirement age
ry <- 2017 - opt_age   # retirement year
domcv10 <- domcv6      # copy of the data


# subset for active wells
domcv10 <- domcv10[which(domcv10@data$year >= ry), ] 
  
# water column heights
# negative water column heights do not affect optimization
# these are just wells that are very far gone
wch <- domcv10@data$bot - domcv10@data$Ensemble
  
# pump separation distance (ft) from top of water column
ps <- opt_dist * wch
  
# pump location (ft below land surface)
pl <- domcv9@data$Ensemble + ps
  
# put into dataframe and compute well failures
domcv10@data <- domcv10@data %>% 
  mutate(pl = pl,
         dry = ifelse(max_gw >= pl, TRUE, FALSE))

# # count dry and wet wells in townships
# blm_dry <- over(domcv10[domcv10@data$dry == T, ], blmcv) %>% count(PLSSID)
# blm_wet <- over(domcv10[domcv10@data$dry == F, ], blmcv) %>% count(PLSSID)
# blm_frp <- left_join(blm_wet, blm_dry, by = "PLSSID") %>% 
#   mutate(total = n.x + n.y,  # total count of wells 
#          frp = (n.y / total) * 100,  # calc failure ratios
#          fc = paste0("(", n.y, "/", total," = ", round(frp, 2), "%)"),
#          fc = str_replace_all(fc, "(NA/NA = NA%)", "NA"))
# 
# blmcv@data <- left_join(blmcv@data, blm_frp, by = "PLSSID")
# 
# # count dry and wet wells in GSAs
# gsa_dry <- over(domcv10[domcv10@data$dry == T, ], gsacv) %>% count(GSA.Name)
# gsa_wet <- over(domcv10[domcv10@data$dry == F, ], gsacv) %>% count(GSA.Name)
# gsa_frp <- left_join(gsa_wet, gsa_dry, by = "GSA.Name") %>% 
#   mutate(total = n.x + n.y,  # total count of wells 
#          frp = (n.y / total) * 100,  # calc failure ratios
#          fc = paste0("(", n.y, "/", total," = ", round(frp, 2), "%)"),
#          fc = str_replace_all(fc, "(NA/NA = NA%)", "NA"))
# 
# gsacv@data <- left_join(gsacv@data, gsa_frp, by = "GSA.Name")
# 
# # count dry and wet wells in b118 
# b118_dry <- over(domcv10[domcv10@data$dry == T, ], b118cv) %>% count(Basin_Subb)
# b118_wet <- over(domcv10[domcv10@data$dry == F, ], b118cv) %>% count(Basin_Subb)
# b118_frp <- left_join(b118_wet, b118_dry, by = "Basin_Subb") %>% 
#   mutate(total = n.x + n.y,  # total count of wells 
#          frp = (n.y / total) * 100,  # calc failure ratios
#          fc = paste0("(", n.y, "/", total," = ", round(frp, 2), "%)"),
#          fc = str_replace_all(fc, "(NA/NA = NA%)", "NA"))
# 
# b118cv@data <- left_join(b118cv@data, b118_frp, by = "Basin_Subb")

# GSAs
gsa_dry <- poly.counts(domcv10[domcv10$dry == T, ], gsacv)
gsa_wet <- poly.counts(domcv10[domcv10$dry == F, ], gsacv)
gsa_frp <- gsa_dry / (gsa_dry + gsa_wet)

gsacv@data$frp <- gsa_frp * 100
gsacv@data$fc <- paste0("(",gsa_dry,"/",(gsa_dry+gsa_wet)," = ",round(gsa_frp*100,2),"%)")
gsacv@data$dry <- gsa_dry

# B118 
b118_dry <- poly.counts(domcv10[domcv10$dry == T, ], b118cv)
b118_wet <- poly.counts(domcv10[domcv10$dry == F, ], b118cv)
b118_frp <- b118_dry / (b118_dry + b118_wet)

b118cv@data$frp <- b118_frp * 100
b118cv@data$fc <- paste0("(",b118_dry,"/",(b118_dry+b118_wet)," = ",round(b118_frp*100,2),"%)")
b118cv@data$dry <- b118_dry

# Township
blm_dry <- poly.counts(domcv10[domcv10$dry == T, ], blmcv)
blm_wet <- poly.counts(domcv10[domcv10$dry == F, ], blmcv)
blm_frp <- blm_dry / (blm_dry + blm_wet)

blmcv@data$frp <- blm_frp * 100
blmcv@data$fc <- paste0("(",blm_dry,"/",(blm_dry+blm_wet)," = ",round(blm_frp*100,2),"%)")
blmcv@data$dry <- blm_dry
```

***  

# Results

### Static Maps

Predictions of domestic well failure for the 2012-2016 drought over different areal units.

```{r}
library(sf)
library(ggplot2)

# convert to sf
blmcvsf  <- st_as_sf(blmcv)
gsacvsf  <- st_as_sf(gsacv)
b118cvsf <- st_as_sf(b118cv)
cvsf     <- st_as_sf(cv)

# plot townships
p1 <- ggplot() +
  geom_sf(data = blmcvsf, aes(fill = log(dry))) +
  geom_sf(data = cvsf, alpha = 0.01, lwd = .8, color = "red") +
  scale_fill_viridis_c("% Failure") +
  coord_sf(crs = st_crs(102003)) +
  labs(#title    = "2012-2016 Domestic Well Failures",
       #subtitle = "Townships",
       y = "Latitude", x = "Longitude") +
  theme_bw()

# plot GSAs
p2 <- ggplot() +
  geom_sf(data = gsacvsf, aes(fill = dry)) +
  geom_sf(data = cvsf, alpha = 0.01, lwd = .8, color = "red") +
  scale_fill_viridis_c("% Failure") +
  coord_sf(crs = st_crs(102003)) +
  labs(#title    = "2012-2016 Domestic Well Failures",
       #subtitle = "Groundwater Sustainability Agencies",
       y = "Latitude", x = "Longitude") +
  theme_bw()

# plot Bulltein 118 subbasins
p3 <- ggplot() +
  geom_sf(data = b118cvsf, aes(fill = dry)) +
  geom_sf(data = cvsf, alpha = 0.01, lwd = .8, color = "red") +
  scale_fill_viridis_c("% Failure") +
  coord_sf(crs = st_crs(102003)) +
  labs(#title    = "2012-2016 Domestic Well Failures",
       #subtitle = "Bulletin 118 Subbasins",
       y = "Latitude", x = "Longitude") +
  theme_bw()

# save
# ggsave(p1, file = "p_blm.png", dpi = 300, height = 10, width = 7)
# ggsave(p2, file = "p_gsa.png", dpi = 300, height = 10, width = 7)
# ggsave(p3, file = "p_118.png", dpi = 300, height = 10, width = 7)

```

#### Township
```{r}
p1 
```

***  

#### Groundwater Sustainability Agency
```{r}
p2
```

***  

#### Bulletin 118 Subbasin
```{r}
p3
```


***  

### Interactive Maps

Predictions for the 2012-2016 drought.

```{r}
library(leaflet)
library(colormap)

# transform polygons to lat/lon for leaflet
b118cvsf <- b118cvsf %>% st_transform(crs = "+init=epsg:4326") 
gsacvsf  <- gsacvsf  %>% st_transform(crs = "+init=epsg:4326") 
blmcvsf  <- blmcvsf  %>% st_transform(crs = "+init=epsg:4326") 

# transform points to ll for leaflet
domcv10ll <- spTransform(domcv10, crs("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0") )

# round well bottom measurements
domcv10ll$bot <- round(domcv10ll$bot, 2)

# icons for points
icons <- icons(
  
)

# function to get colors
getColor <- function(d) {
  sapply(d$dry, function(dry) {
    ifelse(dry == TRUE, "red", "blue") # red if fail, blue if not
  })
}

# make marker list
icons <- awesomeIcons(
  icon = 'f041',
  iconColor = 'black',
  library = 'fa',
  markerColor = getColor(domcv10ll@data)
)


```


#### Township
```{r}
pal <- colorBin(palette = colormap(colormaps$viridis, nshades = 10),
                domain = blmcvsf$frp, bins = seq(0,100,10))

lng <- mean(blmcv@bbox[1, ])
lat <- mean(blmcv@bbox[2, ])

blmcvsf %>% 
  # only show townships with 30 or more wells to begin with
  #mutate(frp = ifelse(nwells >= 30, frp, NA)) %>% 
  leaflet(width = "100%") %>% 
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(label = ~ paste(TWNSHPLAB, fc),
              color = ~ pal(frp),
              stroke = FALSE,
              fillOpacity = 0.7) %>% 
  addPolygons(label = ~ paste(TWNSHPLAB, fc),
              color = ~ pal(frp),
              stroke = FALSE,
              fillOpacity = 0.7)
  addLegend("bottomright", 
            pal = pal, 
            values = ~ frp,
            title = "% Failure",
            opacity = 1,
            labFormat = function(type, cuts, p) {
              n = length(cuts)
              paste0(cuts[-n], " &ndash; ", cuts[-1])
            }
            )

```

***  

#### Groundwater Sustainability Agency
```{r}
pal <- colorBin(palette = colormap(colormaps$viridis, nshades = 10),
                domain = gsacvsf$frp, bins = seq(0,100,10))

gsacvsf %>% 
  leaflet(width = "100%") %>% 
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(label = ~ paste(as.character(GSA.Name), fc),
              color = ~ pal(frp),
              stroke = FALSE,
              fillOpacity = 0.7,
              group = "GSAs") %>% 
  addAwesomeMarkers(lng = domcv10ll@coords[, 1],
             lat = domcv10ll@coords[, 2],
             popup = paste("Well ID:", domcv10ll$WCRNumber,"<br>",
                           "(", domcv10ll$lon, "N", domcv10ll$lat, "W)", "<br>",
                           "Bottom of Perforated Interval:", domcv10ll$bot, "ft.", "<br>",
                           "Dry:", domcv10ll$dry),
             icon = icons,
             group = "Wells",
             clusterOptions = markerClusterOptions()) %>%
  addLayersControl(overlayGroups = c("GSAs", "Wells"), 
                   position = "topleft") %>%
  hideGroup("Wells") %>%
  addLegend("bottomright", 
            pal = pal, 
            values = ~ frp,
            title = "% Failure",
            opacity = 1,
            labFormat = function(type, cuts, p) {
              n = length(cuts)
              paste0(cuts[-n], " &ndash; ", cuts[-1])
            }
            )
```

***  

#### Bulletin 118 Subbasin
```{r}
pal <- colorBin(palette = colormap(colormaps$viridis, nshades = 10),
                domain = b118cvsf$frp, bins = seq(0,100,10))

b118cvsf %>% 
  leaflet(width = "100%") %>% 
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(label = ~ paste(as.character(Subbasin_N), fc),
              color = ~ pal(frp),
              stroke = FALSE,
              fillOpacity = 0.7) %>% 
  addLegend("bottomright", 
            pal = pal, 
            values = ~ frp,
            title = "% Failure",
            opacity = 1,
            labFormat = function(type, cuts, p) {
              n = length(cuts)
              paste0(cuts[-n], " &ndash; ", cuts[-1])
            }
            )

```

