---
title: "Untitled"
output: html_document
---

Before running this script, run `06_calibration_herve.Rmd`.  

The sole purpose of this script it to export data for:  

* Leo : GEOJSON shapefiles to build the interactive visualization.  
* Herve : Machine learning  
* Amanda : social/demographic analysis  
* Alvar: economic analysis  


***  


Cleaning and organizing some data.
```{r}
# polygons: select/create relevant columns for visualization
# and give columns more intuitive names
b118cvsf_export <- b118cvsf %>% 
  select(Subbasin_N, frp, dry, wet, geometry) %>% 
  mutate(total_count = dry + wet,
         frp = round(frp, 2)) %>% 
  rename(name = Subbasin_N, fail_perc = frp, 
         dry_count = dry, wet_count = wet)

blmcvsf_export <- blmcvsf %>% 
  select(TWNSHPLAB, frp, dry, wet, geometry) %>% 
  mutate(total_count = dry + wet,
         frp = round(frp, 2)) %>% 
  rename(name = TWNSHPLAB, fail_perc = frp, 
         dry_count = dry, wet_count = wet)

gsacvsf_export <- gsacvsf %>% 
  select(GSA.Name, frp, dry, wet, geometry) %>% 
  mutate(total_count = dry + wet,
         frp = round(frp, 2)) %>% 
  rename(name = GSA.Name, fail_perc = frp, 
         dry_count = dry, wet_count = wet)

# same for points: first convert to sf object
domcv5llsf <- st_as_sf(domcv5ll) %>% st_transform(crs = "+init=epsg:4326")

domcv5llsf_export <- domcv5llsf %>% 
  select(WCRNumber, bot, dry, geometry) %>% 
  rename(Name = WCRNumber, bottom = bot)
```

Write `.geojson` files for Leo's Mapbox project. 
```{r}
library(geojsonio) # for writing geojson files

geojson_write(b118cvsf_export, geometry = "polygon", group = "name", file = "b118_poly.geojson")
geojson_write(blmcvsf_export, geometry = "polygon", group = "name", file = "blm_poly.geojson")
geojson_write(gsacvsf_export, geometry = "polygon", group = "name", file = "gsa_poly.geojson")
geojson_write(domcv5llsf_export, geometry = "point", group = "name", file = "well_pt.geojson")
```

Write `.shp` files for Amanda's social analysis.  
```{r}
st_write(b118cvsf_export, here("data","amanda_social","b118.shp"))
st_write(blmcvsf_export, here("data","amanda_social","blm.shp"))
st_write(gsacvsf_export, here("data","amanda_social","gsa.shp"))
st_write(domcv5llsf_export %>% select(-bottom), 
         here("data","amanda_social","wells.shp"))
```


Now run `08_make_machine_learning.Rmd` to create data for the machine learning export.  

Write `.shp` files for Herve's machine learning.  

Herve still needs to add:  

* PRISM data  
```{r}

```







