---
title: "Make Drought Scenarios"
output: 
  html_document:
    theme: cosmo
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
    highlight: "pygments"
date: 2018-07-03
author: Rich Pauloo
---

The purpose of this script is to make drought scenarios informed by data from the past (2012-2016) drought.

```{r, echo = FALSE, warning = FALSE, message=FALSE}
library(knitr)
opts_chunk$set(
  fig.width  = 7,
  fig.height = 7,
  collapse   = TRUE,
  message    = FALSE,
  error      = FALSE,
  warning    = FALSE,
  cache      = TRUE
)
```

Load up predicitons of groundwater level.
```{r}
library(raster)
library(tidyverse)
library(here)

# list of gw level interpolations
ml <- read_rds("ml.rds") 

# central valley
cv <- read_rds("cv.rds")

# domestic wells
domcv10 <- read_rds("domcv10.rds")
domcv10 <- spTransform(domcv10, crs(cv))

# extract ensemble prediction
e <- lapply(1:12, function(x){ml[[x]]$Ensemble})
```

```{r}
# pre-drought
pd <- e[[1]]
  
# drought
d <- mean(e[[2]], e[[3]],e[[4]], e[[5]],e[[6]], e[[7]],e[[8]], e[[9]],e[[10]])

# initial drought: spring 2011 and sprign 2012 GW levels
di <- mean(e[[1]], e[[2]])

# final drought: spring and fall 2016 GW levels
df <- mean(e[[9]], e[[10]])
```

Many ways to define the change in GW level over the drought, but we use:  

$\Delta H = H_f - H_i$  

Where:  

$\Delta H$ is the change in groundwater level over the 2012-2016 drought  
$H_f$ is the mean groundwater level of Spring and Fall 2016  
$H_i$ is the mean groundwater level of Spring 2011 and Spring 2012  

We also remove 5% of the outliers fromt he high and low ends of the distribution, which incidentally coincide with areas of low to zero domestic well occurence.  
```{r}
diff <- (df - di)
# diff <- (d - pd)
# diff <- (d- di)


# find the 2.5% and 97.5% quantiles
q <- quantile(na.omit(diff@data@values), probs = c(0.025, 0.975))

# omit these and replot to see main trend
diff2 <- diff
diff2@data@values[diff2@data@values < q[1] | diff@data@values > q[2]] <- NA
```

Visualize.
```{r}
# map and histogram 
diff2 %>% plot()
diff2 %>% hist()
```

Visualize.
```{r}
png(filename = "diff.png", height = 800, width = 800, pointsize = 20)
  par(mfrow=c(2,2))
  
  # plot 1 shows raw data
  plot(diff, main = "Drought - Pre-Drought")
  plot(cv, add = TRUE)
  
  # plot 2 shows omitted extreme values
  plot(diff2, main = "Without Outliers")
  plot(cv, add = TRUE)
  
  # plot of raw data histogram
  hist(diff, main = "", xlab = "GW Depth Below Land Surface (ft.)")
  
  # plot of omitted histogram
  hist(diff2, main = "", xlab = "GW Depth Below Land Surface (ft.)")

dev.off()


png(filename = "diff2.png", height = 600, width = 800, pointsize = 20)
  
  par(mfrow = c(1,2))
  
  # plot 1 shows omitted extreme values
  plot(diff2, main = "Without Outliers")
  plot(cv, add = TRUE)
  
  #plot 2 shows domestic well locations
  plot(diff2, main = "Domestic Well Locations")
  plot(cv, add = TRUE)
  plot(domcv10, pch = 19, cex = 0.1, add = TRUE)
  
dev.off()
```

Scale drought for 1,2,4 years
```{r}
drought_1_yr <- drought_2_yr <- drought_3_yr <- drought_4_yr <- diff2

drought_1_yr@data@values <- diff2@data@values * 0.25
drought_2_yr@data@values <- diff2@data@values * 0.50
drought_3_yr@data@values <- diff2@data@values * 0.75
drought_4_yr@data@values <- diff2@data@values * 1

# visualize
drought_scenarios <- stack(drought_1_yr, drought_2_yr, 
                           drought_3_yr, drought_4_yr)
names(drought_scenarios) <- c("d1", "d2", "d3", "d4")
spplot(drought_scenarios)

# export
write_rds(drought_scenarios, "drought_scenarios.rds")
```

Initial conditions kfor future drougths moving forward == mean(spring and fall 2017 GWL)
```{r}
# mean of spring and fall 2017 Groundwater level is t0 for future droughts
sp_fa_2017_gwl <- mean(e[[11]], e[[12]]) 

# visualize
sp_fa_2017_gwl %>% plot()

# export
write_rds(sp_fa_2017_gwl, "sp_fa_2017_gwl.rds")
```


