---
title: "Density Calculations"
output: 
  html_document:
    theme: cosmo
    toc: TRUE
    toc_float: TRUE
    toc_depth: 2
    code_folding: hide
    highlight: "pygments"
date: 2018-08-09
author: Rich Pauloo
---

With the output from `06_calibration_herve_alvar_graham_calib_TS.Rmd`, I now make density plots of observed and predicted well failures and compare.  


```{r, echo = FALSE, warning = FALSE, message=FALSE}
library(knitr)
opts_chunk$set(
  fig.width  = 7,
  fig.height = 7,
  collapse   = TRUE,
  message    = FALSE,
  error      = FALSE,
  warning    = FALSE,
  cache      = TRUE
)
```

# Data Setup

Packages used.
```{r}
library(here)      # for system agnostic file paths
library(GISTools)  # counting points in polygons with poly.counts()
library(raster)    # for raster objects
library(tidyverse) # general purpose data science toolkit
library(sp)        # spatial objects
library(sf)
```

# Data
```{r}
# mercator projection
merc <- crs("+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 
            +ellps=WGS84 +datum=WGS84 +units=m +no_defs")

# read central valley alluvial basin boundary and transform to mercator
cv <- shapefile(here("data","spatial","central_valley_alluvial_boundary",
                     "Alluvial_Bnd.shp"))
cv <- spTransform(cv, merc)

# observed dry wells
dw <- shapefile(here("data", "dry_wells", "cleaned_dw", "dwcv.shp"))
dw <- spTransform(dw, merc)

# predicted dry wells
pdw <- read_rds(here("code","domcv5ll_dry.rds"))
pdw <- spTransform(pdw, merc)
# pdw_1 <- read_rds(here("code","domcv5ll_dry_1.rds"))
# pdw_1 <- spTransform(pdw_1, merc)

# wells dry at t0
# dry_t0 <- read_rds("dry_at_t0.rds")
# dry_t0 <- spTransform(dry_t0, merc)
# nrow(dry_t0)
```

Subset to study area.
```{r}
b_trim_fill <- read_rds("b_trim_fill.rds")
b_trim_fill <- spTransform(b_trim_fill, merc)
dw <- dw[b_trim_fill, ]
pdw <- pdw[b_trim_fill, ]
```


Compare obsered and predicted dry wells.
```{r}
plot(cv)
plot(dw, pch = 19, cex = 0.01, col = "blue", add=T)
plot(pdw, pch = 19, cex = 0.01, col = "red", add=T)
```

Make pretty
```{r}
# dwsf <-  st_as_sf(dw)
# pdw$geometry <- NULL
# pdwsf <- st_as_sf(pdw)
# cvsf <- st_as_sf(cv)
# 
# p1 <- ggplot() +
#   geom_sf(data = cvsf) +
#   geom_sf(data = dwsf, alpha = 0.4, size = 0.75, color = "red") +
#   labs(title = "Observed Dry Wells",
#        subtitle = paste0("2012-2016 (n = ", nrow(dwsf) ,")")) +
#   theme_minimal()
# 
# p2 <- ggplot() +
#   geom_sf(data = cvsf) +
#   geom_sf(data = pdwsf, alpha = 0.4, size = 0.75, color = "blue") +
#   labs(title = "Predicted Dry Wells",
#        subtitle = paste0("2012-2016 (n = ", nrow(pdwsf) ,")")) +
#   theme_minimal()

# p3 <- cowplot::plot_grid(p1, p2, align = "h")
# ggplot2::ggsave(p3, 
#                 filename = "pred_obs.png", dpi = 300, height = 7, width = 11)
```

```{r}
# plot(cv)
# plot(dry_t0, pch = 19, cex = 0.01, col = "green", add=T)
```


```{r}
# plot(cv)
# plot(dw, pch = 19, cex = 0.01, col = "blue", add=T)
# plot(pdw_1, pch = 19, cex = 0.01, col = "red", add=T)
```

# Point Pattern Analysis: Kernel Density

```{r}
library(spatstat) # for kernel density
library(maptools) # to create owin objects

# make a `ppp` object: points and window of study region

# window: object class "owin"
cv_owin <- maptools::as.owin.SpatialPolygons(cv)

# points 
dw_pts  <- coordinates(dw)
pdw_pts <- coordinates(pdw)
#dry_t0_pts <- coordinates(dry_t0)

# create ppp objects
dw_ppp  <- ppp(dw_pts[, 1],  dw_pts[, 2],  window = cv_owin)
pdw_ppp <- ppp(pdw_pts[, 1], pdw_pts[, 2], window = cv_owin)
#dry_t0_ppp <- ppp(dry_t0_pts[, 1], dry_t0_pts[, 2], window = cv_owin)
```

```{r}
# compute densities: first 
lambda <- nrow(dw) / gArea(cv)
bw <- 0.15/sqrt(5*lambda) # bandwidth -> same as bw.stoyan(dw_ppp)

# calculate gaussian kernel density
dw_dens_1  <- density(pdw_ppp,
                      sigma = bw, adjust = 2)

# visualize
plot(log(dw_dens_1), main = "Observed : Gaussian kernel density")
contour(log(dw_dens_1), add=TRUE)

# function to calculate bandwidth
calc_bandwidth <- function(p, a){
  lambda <- nrow(p)/gArea(a) # p: spatial points; a: polygon. in sq km
  bw <- 0.15/sqrt(5*lambda)
  return(bw)
}
```

Experiments in scaling
```{r}
library(BBmisc) # for scaling matrices

# scale to avoid negative numbers 
dw_dens_1$v[dw_dens_1$v <= 0] <- 0

dw_dens_1$v <- normalize(dw_dens_1$v, 
                          method = "range", 
                          range = c(10,10e8), 
                          on.constant = "quiet")

plot(log(dw_dens_1))
plot(dw_dens_1)

dw_dens_2 <- dw_dens_3 <- dw_dens_1

dw_dens_2$v <- normalize(dw_dens_1$v, 
                          method = "range", 
                          range = c(0,1), 
                          on.constant = "quiet")
dw_dens_3$v <- normalize(log(dw_dens_1$v), 
                          method = "range", 
                          range = c(0,1), 
                          on.constant = "quiet")
plot(dw_dens_2)
plot(dw_dens_3)
temp <- (dw_dens_2 + dw_dens_3) / 2
plot(temp)
```


After selecting the proper bandwidth, we see that the intensity of the point process varies on a log scale. We log-transform the kernel density estimates, and normalize negative values from 0-1 and generate probability maps for the observed and predicted point patterns.
```{r}
# max value in the non-log transformed density estimation
dw_dens_2   <- dw_dens_1
dw_dens_2$v <- log(dw_dens_2$v)
dw_dens_2$v <- normalize(dw_dens_2$v, 
                         method = "range", 
                         range = c(0, 1), 
                         on.constant = "quiet")

dw_dens_2 %>% hist() # ensure data scales

dw_dens_2$v[is.nan(dw_dens_2$v)] <- 0 # -Inf -> 0

plot(cv)
plot(dw_dens_2, main = "Observed - Gaussian kernel density",add=T)

# function to scale data
calc_scaled_density <- function(d){
  d$v[d$v <= 0] <- 0
  
  d$v <- normalize(d$v, 
                   method = "range", 
                   range = c(10,10e8), 
                   on.constant = "quiet")
  
  d_2 <- d_3 <- d
  
  d_2$v <- normalize(d$v, 
                     method = "range", 
                     range = c(0,1), 
                     on.constant = "quiet")
  d_3$v <- normalize(log(d$v), 
                     method = "range", 
                     range = c(0,1), 
                     on.constant = "quiet")
  f <- (d_2 + d_3) / 2
  return(f)
}
```

Apply functions for bandwidth and log normalized probability to observed and predicted dry well kernel densities of point process intensity.
```{r}
# compute bandwidth for observed and predicted point patterns
obw <- calc_bandwidth(dw, cv)
pbw <- calc_bandwidth(pdw, cv)  

# calculate density
od <- density(dw_ppp,  sigma = obw, adjust = 2)  
pd <- density(pdw_ppp, sigma = pbw, adjust = 2)

# scaling values to between 0-1
ods <- calc_scaled_density(od)
pds <- calc_scaled_density(pd)
```

```{r}
# observed kernel density
plot(ods)
plot(cv, add=T)

# predicted kernel density
plot(pds)
plot(cv, add=T)

# difference
plot(pds - ods)
plot(cv, add=T)
```


```{r}
# histogram of difference
data.frame(x = (pds-ods)$v %>% as.vector()) %>% 
  ggplot() +
  #geom_histogram(aes(x)) +
  geom_density(aes(x)) + 
  geom_vline(aes(xintercept = 0.65)) + 
  geom_vline(aes(xintercept = -0.65))
```

The histogram of differences between kernel density estimations looks good, but the kernel density plots themselves, when compared to the maps of point patterns, look off.  

- review kernel density estimation and different bandwidths  
- plot_grid: p1 = observed point pattern, p2 = density of obs pp, p3 = predicted pp, p4 = density of prediction.



# Rasterize and calculate stats

```{r}
s <- 9656.064 # six miles in meters

dwr  <- raster(ods)
crs(dwr) <- crs(dw)
xs <- s/ res(dwr)[1] # x scaling
ys <- s/ res(dwr)[2] # y scaling
dwr <- aggregate(dwr, c(xs, ys))
#dwr@data@values <- dwr@data@values *1000000 * sqrt(100)*2 # m2 -> km2 -> 100km2
spplot(dwr)

pdwr <- raster(pds)
crs(pdwr) <- crs(dw)
xs <- s/ res(pdwr)[1] # x scaling
ys <- s/ res(pdwr)[2] # y scaling
pdwr <- aggregate(pdwr, c(xs, ys))
#pdwr@data@values <- pdwr@data@values *1000000 * sqrt(100)*2 # m2 -> km2 -> 100km2
spplot(pdwr)

# difference between predicted and observed failure density
abs_diff <- abs(pdwr - dwr)
rel_diff <- pdwr - dwr
```

Compare predicted and observed. 
```{r}
# make into data frames
dwr_df  <- as.data.frame(dwr, xy = TRUE)
pdwr_df <- as.data.frame(pdwr, xy = TRUE)

# cv in points for ggplot
cv@data$id = rownames(cv@data)
cv_pts = fortify(cv, region="id")
cv_df = left_join(cv_pts, cv@data, by="id")

# palettes, labels
lab  <- c("0.00 to 0.20","0.20 to 0.40","0.40 to 0.60","0.60 to 0.80", "0.80 to 1.00")

dwr_df$bin  <- base::cut(dwr_df$layer,  breaks = seq(0, 1, 0.2), labels = lab)
pdwr_df$bin <- base::cut(pdwr_df$layer, breaks = seq(0, 1, 0.2), labels = lab)

library(colormap)
pal  <- colormap(colormaps$jet, nshades = 10, alpha = .7)[6:10]
pal[1] <- "#ffffff"

# observed
p_obs_map <- dwr_df %>% 
  ggplot() +
  geom_raster(aes(x,y, fill = bin)) + 
  geom_path(data = cv_df, aes(long, lat, group = group)) +
  coord_fixed(1.1) +
  theme_void() +
  # labs(fill = "Probability",
  #      title = "Observed Log Weighted Kernel Density Estimate",
  #      subtitle = "Spring 2012 - Fall 2016") +
  scale_fill_manual(values = pal,
                    breaks = rev(lab),
                    labels = rev(lab)) +
  guides(fill = FALSE)
  
# predicted
p_pred_map <- pdwr_df %>% 
  ggplot() +
  geom_raster(aes(x,y, fill = bin)) + 
  geom_path(data = cv_df, aes(long, lat, group = group)) +
  coord_fixed(1.1) +
  theme_void() +
  # labs(fill = "Probability",
  #      title = "Predicted Log Weighted Kernel Density Estimate",
  #      subtitle = "Spring 2012 - Fall 2016") +
  scale_fill_manual(values = pal,
                    breaks = rev(lab),
                    labels = rev(lab))+
  guides(fill = FALSE)

# histograms
p_obs_pred_dens <- rbind.data.frame(dwr_df %>% mutate(type = "Observed"), 
                 pdwr_df %>% mutate(type = "Predicted")) %>% 
  ggplot() +
  geom_density(aes(x= layer, fill = type), alpha = 0.3) +
  theme_minimal() +
  labs(#title = "Observed and Predicted Kernel Density Estimates",
       #subtitle = "Spring 2012 - Fall 2016",
       x = "Kernel Density Estimate", y = "Density",
       fill = "") +
  theme(legend.position = "bottom")

# top <- cowplot::plot_grid(p_obs_map, p_pred_map)

# p_kernel_grid <- cowplot::plot_grid(top, p_obs_pred_dens, align = "v", axis = "l", ncol = 1)
#ggsave(p_kernel_grid, filename = "p_kernel_grid.png", height = 7, width = 11, dpi = 300)

# ggplot2::ggsave(top, filename="obs_pred_dens.png", dpi = 300, height = 7, width = 11)
```

Figure for publication. pred obs point patterns
```{r}
p1 <- ggplot() +
  geom_path(data = cv_df, aes(long, lat, group = group)) +
  geom_point(data = as.data.frame(dw_pts), aes(coords.x1,coords.x2), alpha = 0.4, size = 0.75, color = "red") +
  coord_fixed(1.1) +
  theme_void()

p2 <- ggplot() +
  geom_path(data = cv_df, aes(long, lat, group = group)) +
  geom_point(data = as.data.frame(pdw_pts), aes(dom.lon,dom.lat), alpha = 0.4, size = 0.75, color = "blue") +
  coord_fixed(1.1) +
  theme_void()

# p3 <- cowplot::plot_grid(p1, p2, align = "h")


po <- cowplot::plot_grid(p1, p_obs_map, p2, p_pred_map, labels = c("(A)","(B)","(C)","(D)"))

# ggsave(po, filename = "p_o_kernels.pdf", height = 10.5, width = 7.5, device = cairo_pdf)
legend_for_ai <- pdwr_df %>% 
  ggplot() +
  geom_raster(aes(x,y, fill = bin)) + 
  geom_path(data = cv_df, aes(long, lat, group = group)) +
  coord_fixed(1.1) +
  theme_void() +
  # labs(fill = "Probability",
  #      title = "Predicted Log Weighted Kernel Density Estimate",
  #      subtitle = "Spring 2012 - Fall 2016") +
  scale_fill_manual(values = pal,
                    breaks = rev(lab),
                    labels = rev(lab))
# ggsave(legend_for_ai, filename = "legend_for_ai.png", dpi = 300, height = 7, width = 7)

```


Compare difference of rasters.
```{r}
library(colormap)
diff_df <- as.data.frame(abs_diff, xy = TRUE) 
rel_df <- as.data.frame(rel_diff, xy = TRUE) 

lab  <- c("0.00 to 0.20","0.21 to 0.40","0.41 to 0.60","0.61 to 0.80", "0.81 to 1.00")
lab2 <- c("-0.81 to -1.00","-0.61 to -0.80","-0.41 to -0.60","-0.21 to -0.40", "0.00 to -0.20",
          "0.00 to 0.20","0.21 to 0.40","0.41 to 0.60","0.61 to 0.80", "0.81 to 1.00")

diff_df$bin  <- base::cut(diff_df$layer, breaks = seq(0, 1, 0.2), labels = lab)
rel_df$bin2 <- base::cut(rel_df$layer, breaks = seq(-1, 1, 0.2), labels = lab2)

pal  <- colormap(colormaps$jet, nshades =10, alpha = .7)[6:10]
pal[1] <- "#ffffff"
pal2 <- rev(colormap(colormaps$jet, nshades =10, alpha = .7))
pal2[c(5,6)] <- "#ffffff"

# absolute differene map comparing predicted and observed kernel density estimates
p_abs_map <- diff_df %>% 
  ggplot(aes(x,y)) +
  geom_raster(aes(x,y, fill = bin)) +
  geom_path(data = cv_df, aes(long, lat, group = group)) +
  coord_fixed(1.1) +
  theme_void(base_size = 15) +
  labs(#title = "|Predicted - Observed| Kernel Density Estimates",
       #subtitle = "Spring 2012 - Fall 2016",
       fill = "|Difference|") +
  scale_fill_manual(values = pal,
                    breaks = rev(lab),
                    labels = rev(lab))

p_rel_map <- rel_df %>% 
  ggplot(aes(x,y)) +
  geom_raster(aes(x,y, fill = bin2)) +
  geom_path(data = cv_df, aes(long, lat, group = group)) +
  coord_fixed(1.1) +
  theme_void(base_size = 15) +
  labs(#title = "Predicted - Observed Kernel Density Estimates",
       #subtitle = "Spring 2012 - Fall 2016",
       fill = "Difference") +
  scale_fill_manual(values = pal2,
                    breaks = lab2,
                    labels = rev(lab2))


# density plot of absolute difference between predicted and observed
p_abs_dens <- diff_df %>% 
  ggplot(aes(layer)) +
  geom_density(fill = '#619CFF', alpha=0.6) + 
  theme_minimal(base_size = 15) +
  labs(#title = '|Predicted - Observed| Kernel Density Estimates',
       #subtitle = "Spring 2012 - Fall 2016", 
       x = "|Difference|", y = "Density")


# https://stackoverflow.com/questions/4542438/adding-summary-information-to-a-density-plot-created-with-ggplot
q15.9 <- quantile(rel_df$layer, .159, na.rm=T) # 1 Std 68.2%
q84.1 <- quantile(rel_df$layer, .841, na.rm=T)
q2.3  <- quantile(rel_df$layer, .023, na.rm=T) # 2 Std 95.4%
q97.7 <- quantile(rel_df$layer, .977, na.rm=T)
q0.01 <- quantile(rel_df$layer, .001, na.rm=T) # 3 Std 99.8%
q99.9 <- quantile(rel_df$layer, .999, na.rm=T)
meanx <- mean(rel_df$layer, na.rm = T)
medx  <- median(rel_df$layer, na.rm = T)
x.dens  <- density(rel_df$layer, na.rm=T)
df.dens <- data.frame(x=x.dens$x, y=x.dens$y)


p_rel_dens <- ggplot() + 
  geom_area(data = subset(df.dens, x >= q15.9 & x <= q84.1), # 1 Std 68.2%
            aes(x=x,y=y), fill='#619CFF', alpha=0.8) +
  geom_area(data = subset(df.dens, x >= q2.3 & x <= q97.7), # 2 Std 95.4%
            aes(x=x,y=y), fill='#619CFF', alpha=0.6) +
  geom_area(data = subset(df.dens, x >= q0.01 & x <= q99.9), # 3 Std 99.8%
            aes(x=x,y=y), fill='#619CFF', alpha=0.3) +
  geom_vline(xintercept=meanx) +
  geom_vline(xintercept=medx, color='#FFFFFF') +
  geom_density(data = rel_df, aes(x = layer, y = ..density..)) +
  theme_minimal(base_size = 15) +
  labs(#title = "Predicted - Observed Kernel Density Estimates",
       #subtitle = "Spring 2012 - Fall 2016",
       x = "Difference", y = "Density") 
  

p_kernel_diff_grid <- cowplot::plot_grid(p_abs_map, p_abs_dens, p_rel_map, p_rel_dens,
                                    labels = c("(A)","","(B)",""))
# ggsave(p_kernel_diff_grid, filename = "p_kernel_diff_grid.pdf", device = cairo_pdf, width = 11, height = 7)
```



